### 2022-10-29 FDAgroup.Rmd
some new questions?

---
title: "Dynamic Prediction in Functional Concurrent Regression "
output: beamer_presentation
date: "2022-10-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(fcr)
library(face)
library(refund)
library(ggplot2)
library(tidyverse)

set.seed(1025)

df = content
df_train = subset(content, include == 1)
df_test = subset(content, include == 0)
```

## Dynamic prediction

- With observations up to $t_m$, predict outcomes (or probabilities of outcome) after that time point
- Prediction updates with new observations

## Functional Concurrent Regression (FCR)

- For a subject i, we observe a function over t
$$Y_i(t)=f_0(t)+b_i(t)+\epsilon_i(t)$$
- We usually observe $Y_i$ on a series of discrete $t_{ij}$

$$Y_{ij} = f_0(t_{ij})+b_i(t_{ij})+\epsilon_{ij}$$
where $\epsilon_{ij} \sim N(0, \sigma_{\epsilon^2})$.

- Subject-specific random effect

$$b_i(t) = \Sigma_{k=1}^c u_{ik}B_k(t)$$
where $\boldsymbol{u_i}\sim N(0, \Gamma)$

## Dynamic prediction

- Use FPCA to estimate $f_0$, $\Gamma$ and $\sigma_{\epsilon}$

- For a new subject with observations up to $t_m$, estimate its score:

$$\hat{\boldsymbol{u}} = E(\boldsymbol{u}|\boldsymbol{y}) =  \boldsymbol{\hat{\Gamma}B^T}(\boldsymbol{B\hat{\Gamma}B^T}+\hat{\sigma_{\epsilon}}^2\boldsymbol{I_m})$$
With the estimated score, we can predict its outcome in following time points

$$\hat{\boldsymbol{Y}} = \boldsymbol{B}^T\hat{\boldsymbol{u}} $$

## Simulated child growth data

- Predict length-for-age, observed with noise

```{r}
ggplot(df %>% filter(subj%in% 1:10), aes(x=argvals, y=Y, group = subj, col = as.factor(subj)))+
  geom_line(show.legend = F)+
  labs(title = "Length-for-Age for the first 10 subjects",
       x = "time")
```

## FPCA on observed LAZ

```{r}
tnew = sort(unique(df_train$argvals))
df_laz <- df_train %>% select(argvals, subj, Y) %>%
  rename("y" = "Y")
fpca_laz <- face.sparse(df_laz, newdata = df_laz, knots = 15, argvals.new = tnew, calculate.scores = TRUE)

# mean function

par(mfrow = c(2, 2))
plot(tnew, fpca_laz$mu.new, type = "l", 
     xlab = "t", ylab = "f0(t)")
plot(tnew, fpca_laz$eigenfunctions[, 1], type = "l", xlab = "t", ylab = "PC1")
plot(tnew, fpca_laz$eigenfunctions[, 2], type = "l", xlab = "t", ylab = "PC2")
plot(tnew, fpca_laz$eigenfunctions[, 3], type = "l", xlab = "t", ylab = "PC3")

```

## Prediction of new partially observed sample

```{r}
df_laz_test <- df_test %>%  select(argvals, subj, Y) %>%
  rename("y" = "Y")

# observe up to t = 0.3
df_laz_test$y[df_laz_test$argvals > 0.3] <- NA
pred1 <- predict(fpca_laz, newdata = df_laz_test)

# prediction
Bnew <- fpca_laz$eigenfunctions[tnew > 0.3, ]
Yhat <- Bnew %*% t(pred1$rand_eff$scores)
# every column is a new function

Yhat <- data.frame(Yhat)
colnames(Yhat) <- pred1$rand_eff$subj

df_laz_pred <- Yhat %>% 
  mutate(argvals = tnew[tnew > 0.3]) %>% 
  pivot_longer(1:50, names_to = "subj", values_to = "y") %>%
  mutate(type = "pred")

# figure 
bind_rows(df_test %>% 
            select(argvals, subj, Y)%>% 
            rename("y" = "Y") %>% 
            mutate(type = "true", subj = as.factor(subj)), 
          df_laz_pred %>% mutate(subj = as.factor(subj))) %>% 
  filter(subj %in% 151:156) %>% 
  ggplot(aes(x = argvals, y = y, col = type))+
  geom_line()+
  facet_wrap(~subj)+
  labs(title = "Prediction with observation up to t=0.3", 
       x = "time", y = "")

```

## Prediction of new partially observed sample

```{r}
df_laz_test <- df_test %>%  select(argvals, subj, Y) %>%
  rename("y" = "Y")

# observe up to t = 0.3
df_laz_test$y[df_laz_test$argvals > 0.5] <- NA
pred1 <- predict(fpca_laz, newdata = df_laz_test)

# prediction
Bnew <- fpca_laz$eigenfunctions[tnew > 0.5, ]
Yhat <- Bnew %*% t(pred1$rand_eff$scores)
# every column is a new function

Yhat <- data.frame(Yhat)
colnames(Yhat) <- pred1$rand_eff$subj

df_laz_pred <- Yhat %>% 
  mutate(argvals = tnew[tnew > 0.5]) %>% 
  pivot_longer(1:50, names_to = "subj", values_to = "y") %>%
  mutate(type = "pred")

# figure

bind_rows(df_test %>% 
            select(argvals, subj, Y)%>% 
            rename("y" = "Y") %>% 
            mutate(type = "true", subj = as.factor(subj)), 
          df_laz_pred %>% mutate(subj = as.factor(subj))) %>% 
  filter(subj %in% 151:156) %>% 
  ggplot(aes(x = argvals, y = y, col = type))+
  geom_line()+
  facet_wrap(~subj)+
  labs(title = "Prediction with observation up to t=0.5", 
       x = "time", y = "")

```

## Prediction of new partially observed sample

```{r}
df_laz_test <- df_test %>%  select(argvals, subj, Y) %>%
  rename("y" = "Y")

# observe up to t = 0.3
df_laz_test$y[df_laz_test$argvals > 0.8] <- NA
pred1 <- predict(fpca_laz, newdata = df_laz_test)

# prediction
Bnew <- fpca_laz$eigenfunctions[tnew > 0.8, ]
Yhat <- Bnew %*% t(pred1$rand_eff$scores)
# every column is a new function

Yhat <- data.frame(Yhat)
colnames(Yhat) <- pred1$rand_eff$subj

df_laz_pred <- Yhat %>% 
  mutate(argvals = tnew[tnew > 0.8]) %>% 
  pivot_longer(1:50, names_to = "subj", values_to = "y") %>%
  mutate(type = "pred")

# figure

bind_rows(df_test %>% 
            select(argvals, subj, Y)%>% 
            rename("y" = "Y") %>% 
            mutate(type = "true", subj = as.factor(subj)), 
          df_laz_pred %>% mutate(subj = as.factor(subj))) %>% 
  filter(subj %in% 151:156) %>% 
  ggplot(aes(x = argvals, y = y, col = type))+
  geom_line()+
  facet_wrap(~subj)+
  labs(title = "Prediction with observation up to t=0.8", 
       x = "time", y = "")

```

## Next steps

- Establish interval prediction
- ? fPCA on pooled prediction, instead of original functions observed with noise
- Extension to exponential family data