### 2023-03-30 Write-ups/Manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage[]{mathpazo}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}\doublespacing
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Dynamic Prediction of Non-Guassian Outcome with fast Generalized Functional Principal Analysis},
  pdfauthor={Ying Jin; Andrew Leroux},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Dynamic Prediction of Non-Guassian Outcome with fast Generalized
Functional Principal Analysis}
\author{Ying Jin \and Andrew Leroux}
\date{March 30, 2023}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Biomedical investigators are often interested in predicting future
observations of subjects based on their historical data, typically
referred to as dynamic prediction. Traditionally, this type of data has
been modeled using marginal models (generalized estimating equations) or
conditional models (mixed effect models)
(\protect\hyperlink{ref-Laird1982}{Laird and Ware 1982};
\protect\hyperlink{ref-liang1986}{LIANG and ZEGER 1986};
\protect\hyperlink{ref-lindstrom1990}{Lindstrom and Bates 1990};
\protect\hyperlink{ref-davidian2003}{{``Nonlinear models for repeated
measurement data''} 2003}), and predictions are made based on the
correlation between repeated measures from the same subject, and/or
covariates that can be either fixed or time-varying. However, these
methods are limited in terms of flexibility of correlation structure and
the ability to handle out-of-sample prediction. When sample size is
large or the density of repeated measures is high, they also tend to
cause severe computational burden
(\protect\hyperlink{ref-GLMMadaptive}{Rizopoulos 2022}).

To address these problems, one may turn to functional mixed effect
models instead when measures are dense across the domain of measures.
Such methods accommodate more flexible correlation structure by modeling
subject-specific random effects as a function, and non-parametric
smoothing (\protect\hyperlink{ref-Scheipl2014}{Scheipl et al. 2014}) can
be incorporated to speed up the computation, such as spline basis
functions or eigenfunctions from functional principal component analysis
(FPCA). The introduction of basis functions also makes out-of-sample
prediction more straightforward. Instead of estimating subject-specific
random effects of new observations, we can simply estimate
coefficients/loadings on the basis function used for smoothing. Existing
research on dynamic prediction with functional data analysis methods has
been focusing on continuous/Gaussian outcomes, modelling
subject-specific random effects with FPCA
(\protect\hyperlink{ref-chiou2012}{Chiou 2012};
\protect\hyperlink{ref-goldberg2014}{Goldberg et al. 2014};
\protect\hyperlink{ref-shang2017}{Shang 2017}). Kraus
(\protect\hyperlink{ref-kraus2015}{2015}) has used this approach to
predict missing observations in partially observed function tracks, and
Delaigle and Hall (\protect\hyperlink{ref-delaigo2016}{2016}) achieved
similar goals using Markov Chains. While methods mentioned above used
only partial observations for prediction with an intercept-only model,
Leroux et al. (\protect\hyperlink{ref-leroux2016}{2018}) proposed
Functional Concurrent Regression (FCR) framework which can incorporate
the effect of subject-specific predictors.

Unfortunately, fewer papers have focused on its extension to
non-Gaussian data, such as series of binary or count outcomes. Existing
methods also tend to be very computationally intensive. For example,
Chen et al. (\protect\hyperlink{ref-chen2013}{2013}) proposed approaches
to fit marginal functional models that is compatible to multi-level,
generalized outcomes. Goldsmith et al.
(\protect\hyperlink{ref-goldsmith2015}{2015}) established a model
framework that takes into account the fixed effect of time-invariant
covariates, with parameters estimated with Bayesian method in
\emph{Stan}. Gertheiss et al.
(\protect\hyperlink{ref-gertheiss2016}{2016}) identified bias introduced
by directly applying FPCA methods to generalized functions, and proposed
to address this problem using a two-stage, joint estimation strategy.
Linde (\protect\hyperlink{ref-linde2019}{2009}) used an adapted Bayesian
variational algorithm for FPCA of binary and count data. In terms of
implementation, Wrobel et al. (\protect\hyperlink{ref-wrobel2019}{2019})
proposed a fast, efficient way to fit GFPCA on binary data using EM
algorithm, accompanied by the an open source R package \emph{registr}.

In this paper, we aim to develop a fast, scalable method for dynamic
prediction of discrete function tracks based on functional mixed effect
model with FPCA smoothing. Section 2 presents the procedure of the
proposed method. In Section 3, we illustrate the performance and
efficiency of our proposed method in a simulation study. In Section 4,
we apply this method to a real-world dataset. Section 5 presents a
discussion of advantages and limitation of the proposed method.

\hypertarget{method}{%
\section{Method}\label{method}}

The observed data for a single subject \(i\) is (\(t\), \(Y_i(t)\)),
where \(t\) consists of dense, discrete points along the functional
domain, and \(Y_i(t)\) is the non-Gaussian outcome observed at \(t\). We
assume that the outcome \(Y_i(t)\) can be characterized by a latent
continuous function \(\eta_i(t)\). That is, at a specific t, \(Y_i(t)\)
follows a exponential family distribution such that:

\[
g[E(Y_i(t))] = \eta_i(t) = \beta_0(t)+b_i(t)
\]

where g is an appropriate link function, \(\beta_0(t)\) is the
population mean of latent function, and \(b_i(t)\) is a subjects-level
random effect function following a zero-mean Gaussian process.

According to the Karhunen-Loève theorem, the unobserved random effect
function \(b_i(t)\) can be represented as the linear combination of a
infinite set of orthogonal functions:
\(b_i(t)=\sum_{k=1}^{\infty}\xi_{ik}\phi_{k}(t)\). We approximate this
function by a finite set of \(\phi_{k}(t)\), such that
\(b_i(t)=\sum_{k=1}^{K}\xi_{ik}\phi_{k}(t)+\epsilon_i(t)\), where
\(\epsilon_i(t)\) indicates random noise introduced by this
approximation, following a zero-mean Gaussian process. At each specific
point t, \(\epsilon_i(t) \sim N(0, \sigma^2)\).

In practice, this approximation could be implemented through an FPCA
procedure on \(\eta_i(t)\). In this case
\(\phi_{k}(t),\ k \in \{1,...,K\}\) are the first K eigenfunctions that
explain the most variation in \(\eta_i(t)\), and \(\xi_{ik}\) are
subject-specific PC scores (or loadings) on each eigenfunction.
Additionally, \(\xi_{ik}\) are mutually independent over both subject
(\(i\)) and eigenfunctions (\(k\)). That is, each \(\xi_{ik}\) follows
normal distribution \(N(0, \lambda_k)\) where \(\lambda_k\) is the kth
eigenvalues.

\textbf{Fast Generalized FPCA} Based on the problem set up above, we
propose the following algorithm for fast, generalizable implementation
of FPCA (fGFPCA) on the unobserved latent process \(\eta_i(t)\):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Bin the observed outcomes in to small, non-overlapping, equal length
  intervals. We hereafter index the bins by their midpoints \(s\).
\item
  Fit a local Generalized Mixed Model at every bin. Specifically, at bin
  \(s\), we fit \(g[E(Y_i(t))] = \beta_0(s)+b_i(s)\) for all \(t\) in
  bin \(s\). From this series of models we can estimate the individual
  latent functions at every bin:
  \(\hat{\eta}_i(s) = \hat{\beta}_0(s)+\hat{b}_i(s)\).
\item
  Fit FPCA on the estimated latent functions:
  \(\hat{\eta}_i(s) = f_0(s)+\sum_{k=1}^K\xi_{ik}\phi_{k}(s)+\epsilon_i(s)\),
  and obtain estimates of basis functions
  \(\boldsymbol{\Phi} = \{\phi_1(s), ...,\phi_k(s)\}\), eigenvalues
  \(\hat{\lambda}_1...\hat{\lambda}_k\), mean function \(\hat{f}_0(s)\)
  and residual variance \(\hat{\sigma}^2\).
\end{enumerate}

\textbf{Dynamic prediction} Now assume we have a new observations with
partially observed outcome \(Y_{new}(t)\), \(t\leq t_m\). The maximum
observed point is \(t_m\), which is a point along the functional domain
and belongs to bin \(s_m\) on the binned grid from step 1 above. With
components extracted from the FPCA model in step 3, we will be able to
make predictions on the future, unobserved bins. Specifically, with the
partially observed binary function \(Y_{new}(t)\), we will be able to
calculate the maximum likelihood estimate (MLE) of the subject-specific
PC scores \(\hat{\xi}_{ik}\). Then the value of latent functions at
unobserved points can be estimated as
\(\hat{\eta_i}(s)=\hat{f}_0(s)+\sum_{k=1}^K\hat{\xi}_{ik}{\phi}_k(s)\)

Following the algorithm above, predictions of individual latent
functions are made on the binned grid based on partially observed
non-Gaussian functions tracks of new subjects. Since the bins are set up
to be small in length, the binned grid would still be dense. However, in
situations where predictions on the original, un-binned grid is needed,
linear interpolation turns out to be a fast, convenient way with good
performance for prediction at points between bins.

\hypertarget{simulation}{%
\section{Simulation}\label{simulation}}

In this section, we illustrate the predictive performance and
computational efficiency of the proposed method through a simulation
study. We simulated 50 datasets, each with 500 subjects. For every
subject, we generate 1000 binary outcomes \(Y_i(t) \in \{0, 1\}\) across
functional domain \(t \in [0, 1000]\), where the distribution of outcome
is characterized by a continuous latent function. The data generation
mechanism can be expressed as follows:

\[\begin{aligned}
Y_i(t) & \sim Binomial(\frac{exp(\eta_i(t))}{1+exp(\eta_i(t))}) \\
\eta_i(t) &= f_0(t)+ \xi_{i1}\sqrt{2}sin(2\pi t)+\xi_{i2}\sqrt{2}cos(2\pi t)+\xi_{i3}\sqrt{2}sin(4\pi t)+\xi_{i4}\sqrt{2}cos(4\pi t)
\end{aligned}\]

In this simulation, we set \(f_0(t) = 0\). \(\xi_{ik}\) are mutually
independent normal random variables \(\xi_{ik}\sim N(0, \gamma_k)\).
Here we set the values of \(\gamma_k\) to be \(0.5^{k-1}\),
\(k \in (1,..., 4)\). In addition, for simplicity of presentation, we
generate data on a regular grid, which means observations points are
equally distributed across \([0, 1000]\) and are the same for all
subjects.

We use two metrics to evaluate the out-of-sample predictive performance:
integrated squared error (ISE) and
Area-Under-the-Receiver-Operator-Curve (AUC). ISE assess the prediction
accuracy of latent continuous function. It is evaluated on the binned
grid at midpoints of each unobserved bin. If the entire functional
domain has S bins, but we have observations up to the mth bin, then ISE
is defined as
\(\frac{1}{N}\sum_{i=1}^N\sum_{s=s_m+1}^S(\hat{\eta}_i(s)-\eta_i(s))^2\).
The second metric, AUC, focuses on evaluation of prediction of the
binary outcome. Since the binary outcomes are generated on the original,
un-binned grid, we evaluated AUC on this grid and estimated values of
latent functions between bins with linear interpolation.

As a reference method, we compare our method to Generalized Linear Mixed
Models using Adaptive Gaussian Quadrature (GLMMadaptive). This is one of
the fastest existing method developed for dynamic prediction of repeated
generalized outcomes. Just like many mixed models, this method is very
limited in terms of flexibility. For example, the model used for
prediction of our simulated datset would simply be an linear model with
one covariate indicating observation time:
\(g(E(Y_i)) = \beta_0+\beta_1t+b_{i0}+b_{i1}t\). While the flexibility
of this mixed model can be increased using spline functions, the
dimension of spline basis is also restricted by computational ability,
and is unfeasible to implement under the scale of our simulated data or
the complexity of our proposed method.

The average ISE and AUC across all simulation is presented in Table 1,
and Figure 1 presents the predicted function curves of four
randomly-drawn subjects from the first simulation. The prediction is
conditioning on different length of observed track (specifically, with
observations up to t = 200, 400, 600, 800 respectively), and evaluation
is made on equal-length time windows on the unobserved tracks following
the maximum observation time. As Table 1 reveals, our proposed method
(fGFPCA) outperformed GLMMadaptive under every scenario, which was
expected since GLMMadaptive can only accommodate simple model
structures. In addition, fGFPCA also took less computation time. For one
simulated dataset, fGFPCA spent 4.94 minutes on model fitting and
out-of-sample prediction, while GLMMadaptive took 5.11 minutes.

Because of the flexibility of our proposed model framework, the accuracy
of prediction at specific time points would improve with more observed
data. This is revealed in Table 1 as ISE decreases and AUC increases
with maximum observed time (left-to-right), also in Figure 1 as
predicted curves get closer with longer observed track. However the same
tendency is not observed with GLMMadaptive models, as a result of
restricted model flexibility. A linear model that fits well to a
specific part of latent function can fit very badly to the following
parts, especially when the underlying latent function has cyclic
patterns. Therefore, more observations do not necessarily make the model
more predictive.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/Figure-1.pdf}
\caption{Predicted track of four randomly selected subjects from one
simulated dataset. The grey solid line indicates the true latent
continuous function. The dashed lines indicated predicted latent
function tracks, and color indicates different observation time.}
\end{figure}

\begin{table}

\caption{\label{tab:format table}Predictive performance of fGFPCA and GLMM adaptive on the simulated datasets. ISE and AUC are average values across all 50 simulations.}
\centering
\begin{tabular}[t]{llccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{8}{c}{Maximum observed time} \\
\cmidrule(l{3pt}r{3pt}){2-9}
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{fGFPCA} & \multicolumn{4}{c}{GLMMadaptive} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & 200 & 400 & 600 & 800 & 200 & 400 & 600 & 800\\
\midrule
\addlinespace[0.3em]
\multicolumn{9}{l}{\textbf{Prediction time window}}\\
\addlinespace[0.3em]
\multicolumn{9}{l}{\textbf{ISE}}\\
\hspace{1em}\hspace{1em}(200, 400] & 15.29 &  &  &  & 38.20 &  &  & \\
\hspace{1em}\hspace{1em}(400, 600] & 18.68 & 8.53 &  &  & 28.59 & 26.46 &  & \\
\hspace{1em}\hspace{1em}(600, 800] & 22.53 & 5.51 & 1.65 &  & 31.32 & 27.97 & 27.40 & \\
\hspace{1em}\hspace{1em}(800, 1000] & 11.34 & 8.72 & 1.90 & 1.30 & 55.19 & 46.89 & 58.57 & 59.25\\
\addlinespace[0.3em]
\multicolumn{9}{l}{\textbf{AUC}}\\
\hspace{1em}\hspace{1em}(200, 400] & 0.74 &  &  &  & 0.59 &  &  & \\
\hspace{1em}\hspace{1em}(400, 600] & 0.66 & 0.73 &  &  & 0.52 & 0.59 &  & \\
\hspace{1em}\hspace{1em}(600, 800] & 0.71 & 0.79 & 0.80 &  & 0.67 & 0.70 & 0.69 & \\
\hspace{1em}\hspace{1em}(800, 1000] & 0.74 & 0.75 & 0.78 & 0.78 & 0.52 & 0.56 & 0.53 & 0.57\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{data-application}{%
\section{Data application}\label{data-application}}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/Figure_appl-1.pdf}
\caption{Predicted track of four randomly selected subjects from the
daily weigh-in dataset. The grey solid line indicates the true latent
continuous function. The dashed lines indicated predicted latent
function tracks, and color indicates different observation time.}
\end{figure}

\begin{table}

\caption{\label{tab:unnamed-chunk-3}AUC of fGFPCA and GLMM adaptive on the daily weigh-in dataset}
\centering
\begin{tabular}[t]{llccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Maximum observed time} \\
\cmidrule(l{3pt}r{3pt}){2-7}
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{fGFPCA} & \multicolumn{3}{c}{GLMMadaptive} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
  & 100 & 200 & 300 & 100 & 200 & 300\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Prediction time window}}\\
\hspace{1em}(100, 200] & 0.75 &  &  & 0.76 &  & \\
\hspace{1em}(200, 300] & 0.77 & 0.82 &  & 0.76 & 0.80 & \\
\hspace{1em}(300, 400] & 0.76 & 0.81 & 0.84 & 0.74 & 0.78 & 0.82\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\begin{itemize}
\tightlist
\item
  Grid
\item
  Score bias: cannot demonstrate without repeat simulation
\end{itemize}

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-chen2013}{}}%
Chen, H., Wang, Y., Paik, M. cho, and Choi, H. A. (2013), {``A marginal
approach to reduced-rank penalized spline smoothing with application to
multilevel functional data,''} \emph{J Am Stat Assoc.}, 108, 1216--1229.
\url{https://doi.org/10.1080/01621459.2013.826134}.

\leavevmode\vadjust pre{\hypertarget{ref-chiou2012}{}}%
Chiou, J.-M. (2012), {``Dynamical functional prediction and
classification, with application to traffic flow prediction,''}
\emph{The Annals of Applied Statistics}, Institute of Mathematical
Statistics, 6, 1588--1614. \url{https://doi.org/10.1214/12-AOAS595}.

\leavevmode\vadjust pre{\hypertarget{ref-delaigo2016}{}}%
Delaigle, A., and Hall, P. (2016), {``Approximating fragmented
functional data by segments of markov chains,''} \emph{Biometrika}, 103,
779--799. \url{https://doi.org/10.1093/biomet/asw040}.

\leavevmode\vadjust pre{\hypertarget{ref-gertheiss2016}{}}%
Gertheiss, J., Goldsmith, J., and Staicu, A. (2016), {``A note on
modeling sparse exponential-family functional response curves,''}
\emph{Comput Stat Data Anal}, 105, 46--52.
\url{https://doi.org/10.1016/j.csda.2016.07.010}.

\leavevmode\vadjust pre{\hypertarget{ref-goldberg2014}{}}%
Goldberg, Y., Ritov, Y., and Mandelbaum, A. (2014), {``Predicting the
continuation of a function with applications to call center data,''}
\emph{Journal of Statistical Planning and Inference}, 147, 53--65.
https://doi.org/\url{https://doi.org/10.1016/j.jspi.2013.11.006}.

\leavevmode\vadjust pre{\hypertarget{ref-goldsmith2015}{}}%
Goldsmith, J., Zipunnikov, V., and Schrack, J. (2015), {``Generalized
multilevel function-on-scalar regression and principal component
analysis,''} \emph{Biometrics}, 71, 344--53.
\url{https://doi.org/10.1111/biom.12278}.

\leavevmode\vadjust pre{\hypertarget{ref-hall2018}{}}%
Hall, P., Müller, H.-G., and Yao, F. (2008), {``Modelling sparse
generalized longitudinal observations with latent gaussian processes,''}
\emph{Journal of the Royal Statistical Society: Series B (Statistical
Methodology)}, 70, 703--723.
https://doi.org/\url{https://doi.org/10.1111/j.1467-9868.2008.00656.x}.

\leavevmode\vadjust pre{\hypertarget{ref-kraus2015}{}}%
Kraus, D. (2015),
{``\href{http://www.jstor.org/stable/24775309}{Components and completion
of partially observed functional data},''} \emph{Journal of the Royal
Statistical Society. Series B (Statistical Methodology)}, {[}Royal
Statistical Society, Wiley{]}, 77, 777--801.

\leavevmode\vadjust pre{\hypertarget{ref-Laird1982}{}}%
Laird, N. M., and Ware, J. H. (1982),
{``\href{http://www.jstor.org/stable/2529876}{Random-effects models for
longitudinal data},''} \emph{Biometrics}, {[}Wiley, International
Biometric Society{]}, 38, 963--974.

\leavevmode\vadjust pre{\hypertarget{ref-leroux2022}{}}%
Leroux, A., Crainiceanu, C. M., and Wrobel, J. (n.d.). {``Fast
generalized functional principal component analysis.''}

\leavevmode\vadjust pre{\hypertarget{ref-leroux2016}{}}%
Leroux, A., Xiao, L., Crainiceanu, C., and Checkley, W. (2018),
{``Dynamic prediction in functional concurrent regression with an
application to child growth,''} \emph{Statistics in medicine}, 37,
1376--1388.

\leavevmode\vadjust pre{\hypertarget{ref-liang1986}{}}%
LIANG, K.-Y., and ZEGER, S. L. (1986), {``Longitudinal data analysis
using generalized linear models,''} \emph{Biometrika}, 73, 13--22.
\url{https://doi.org/10.1093/biomet/73.1.13}.

\leavevmode\vadjust pre{\hypertarget{ref-linde2019}{}}%
Linde, van der (2009), {``A bayesian latent variable approach to
functional principal components analysi with binary and count data,''}
\emph{A StA Adv Stat Anal}, 307--333.
\url{https://doi.org/10.1007/s10182-009-0113-6}.

\leavevmode\vadjust pre{\hypertarget{ref-lindstrom1990}{}}%
Lindstrom, M. J., and Bates, D. M. (1990),
{``\href{http://www.jstor.org/stable/2532087}{Nonlinear mixed effects
models for repeated measures data},''} \emph{Biometrics}, {[}Wiley,
International Biometric Society{]}, 46, 673--687.

\leavevmode\vadjust pre{\hypertarget{ref-davidian2003}{}}%
{``\href{http://www.jstor.org/stable/1400665}{Nonlinear models for
repeated measurement data: An overview and update}''} (2003),
{[}International Biometric Society, Springer{]}, 8, 387--419.

\leavevmode\vadjust pre{\hypertarget{ref-GLMMadaptive}{}}%
Rizopoulos, D. (2022),
\emph{\href{https://CRAN.R-project.org/package=GLMMadaptive}{GLMMadaptive:
Generalized linear mixed models using adaptive gaussian quadrature}}.

\leavevmode\vadjust pre{\hypertarget{ref-Scheipl2014}{}}%
Scheipl, F., Staicu, A.-M., and Greven, S. (2014), {``Functional
additive mixed models,''} \emph{J Comput Graph Stat}, 24, 447--501.
\url{https://doi.org/10.1080/10618600.2014.901914}.

\leavevmode\vadjust pre{\hypertarget{ref-shang2017}{}}%
Shang, H. L. (2017), {``Functional time series forecasting with dynamic
updating: An application to intraday particulate matter
concentration,''} \emph{Econometrics and Statistics}, 1, 184--200.
https://doi.org/\url{https://doi.org/10.1016/j.ecosta.2016.08.004}.

\leavevmode\vadjust pre{\hypertarget{ref-suresh2017}{}}%
Suresh, K., Taylor, J. M. G., Spratt, D. E., Daignault, S., and
Tsodikov, A. (2017), {``Comparison of joint modeling and landmarking for
dynamic prediction under an illness-death model,''} \emph{Biom J}, 59,
1277--1300. \url{https://doi.org/10.1002/bimj.201600235}.

\leavevmode\vadjust pre{\hypertarget{ref-wrobel2019}{}}%
Wrobel, J., Zipunnikov, V., Schrack, J., and Goldsmith, J. (2019),
{``Registration for exponential family functional data,''}
\emph{Biometrics}, 75, 48--57.
https://doi.org/\url{https://doi.org/10.1111/biom.12963}.

\end{CSLReferences}

\end{document}