### 2023-04-03 Write-ups/Manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage[]{mathpazo}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}\doublespacing
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Dynamic Prediction of Non-Guassian Outcome with fast Generalized Functional Principal Analysis},
  pdfauthor={Ying Jin; Andrew Leroux},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Dynamic Prediction of Non-Guassian Outcome with fast Generalized
Functional Principal Analysis}
\author{Ying Jin \and Andrew Leroux}
\date{April 03, 2023}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Biomedical investigators are often interested in predicting future
observations of subjects based on their historical data, typically
referred to as dynamic prediction. Traditionally, this type of data has
been modeled either marginally, such as generalized estimating
equations, or constitutionally on specific subjects, such as mixed
effect models (\protect\hyperlink{ref-Laird1982}{Laird and Ware 1982};
\protect\hyperlink{ref-liang1986}{LIANG and ZEGER 1986};
\protect\hyperlink{ref-lindstrom1990}{Lindstrom and Bates 1990};
\protect\hyperlink{ref-davidian2003}{{``Nonlinear models for repeated
measurement data''} 2003}). For subject-specific models, predictions are
made based on the correlation between repeated measures from the same
subject and covariates that can be either fixed or varying. However,
most of these methods are based on linear model structure, which is very
limited in terms of flexibility of correlation. When sample size is
large or the density of repeated measures is high, they also tend to
cause severe computational burden
(\protect\hyperlink{ref-GLMMadaptive}{Rizopoulos 2022}). In addition,
with generalized, non-Gaussian outcomes, these methods lack the ability
to handle out-of-sample prediction because there is usually no
close-form solution for subject-specific random effects

To address these problems, one may turn to functional mixed effect
models instead when measures are dense across the domain of measures.
Such methods accommodate more flexible correlation structure by modeling
subject-specific random effects as a function. Non-parametric smoothing
(\protect\hyperlink{ref-Scheipl2014}{Scheipl et al. 2014}) are often
incorporated to speed up the computation, where random effects are
approximated by spline basis functions or eigenfunctions from functional
principal component analysis (FPCA). The introduction of basis functions
also makes out-of-sample prediction more straightforward. Instead of
estimating subject-specific random effects of new observations, we can
simply estimate coefficients/loadings on the basis function used for
smoothing. Existing research on dynamic prediction with functional data
analysis methods has been focusing on continuous/Gaussian outcomes,
usually modelling subject-specific random effects with FPCA
(\protect\hyperlink{ref-chiou2012}{Chiou 2012};
\protect\hyperlink{ref-goldberg2014}{Goldberg et al. 2014};
\protect\hyperlink{ref-shang2017}{Shang 2017}). Kraus
(\protect\hyperlink{ref-kraus2015}{2015}) has used this approach to
predict missing observations in partially observed function tracks, and
Delaigle and Hall (\protect\hyperlink{ref-delaigo2016}{2016}) achieved
similar goals using Markov Chains. While methods mentioned above used
only partial observations for prediction with an intercept-only model,
Leroux et al. (\protect\hyperlink{ref-leroux2016}{2018}) proposed
Functional Concurrent Regression (FCR) framework which can incorporate
the effect of subject-specific predictors.

Unfortunately, fewer papers have focused on its extension to
generalized, non-Gaussian outcomes, such as series of binary or count
outcomes. Existing methods also tend to be very computationally
intensive. For example, Chen et al.
(\protect\hyperlink{ref-chen2013}{2013}) proposed approaches to fit
marginal functional models that is compatible to multi-level,
generalized outcomes. Goldsmith et al.
(\protect\hyperlink{ref-goldsmith2015}{2015}) established a model
framework that takes into account the fixed effect of time-invariant
covariates, with parameters estimated with Bayesian method in
\emph{Stan}. Gertheiss et al.
(\protect\hyperlink{ref-gertheiss2016}{2016}) identified bias introduced
by directly applying FPCA methods to generalized functions, and proposed
to address this problem using a two-stage joint estimation strategy.
Linde (\protect\hyperlink{ref-linde2019}{2009}) used an adapted Bayesian
variational algorithm for FPCA of binary and count data. In terms of
implementation, Wrobel et al. (\protect\hyperlink{ref-wrobel2019}{2019})
proposed a fast, efficient way to fit GFPCA on binary data using EM
algorithm, accompanied by the an open source R package \emph{registr}.

In this paper, we aim to develop a fast, scalable method for dynamic
prediction of generalized functional outcome based on functional mixed
effect model with FPCA smoothing. Specifically, ``Generalized''
indicates that the outcome follows non-Gaussian exponential family
distribution, and the term ``functional'' indicates that repeated
measure are collected densely across the study domain (e.g., time).
Section 2 explains the implementation procedure of the proposed method.
In Section 3, we illustrate the performance and efficiency of our
proposed method in a simulation study. In Section 4, we apply this
method to a real world dataset from a weight loss study. Section 5
presents a discussion of advantages and limitation of the proposed
method.

\hypertarget{method}{%
\section{Method}\label{method}}

The observed data for a single subject \(i\) is (\(t\), \(Y_i(t)\)),
where \(t\) consists of dense, discrete points along the domain of data
collection. In practice, \(t\) is usually a time index along the
duration of study. We hereafter refer to the range of \(t\) as the
``functional domain''. For simplicity, we further assume that outcomes
are collected on a regular grid over the functional domain, meaning that
\(t\) are evenly distributed and does not change across subjects.

\(Y_i(t)\) is the non-Gaussian outcome observed at \(t\). We assume that
the outcome \(Y_i(t)\) can be characterized by a latent continuous
function \(\eta_i(t)\). That is, at a specific t, \(Y_i(t)\) follows a
exponential family distribution such that:

\[
g[E(Y_i(t))] = \eta_i(t) = \beta_0(t)+b_i(t)
\]

where g is an appropriate link function, \(\beta_0(t)\) is the
population mean of latent function, and \(b_i(t)\) is a subjects-level
random effect function following a zero-mean Gaussian process.

According to the Karhunen-Loève theorem, the unobserved random effect
function \(b_i(t)\) can be represented as the linear combination of a
infinite set of orthogonal functions:
\(b_i(t)=\sum_{k=1}^{\infty}\xi_{ik}\phi_{k}(t)\). We approximate this
function by a finite set of \(\phi_{k}(t)\), such that
\(b_i(t)=\sum_{k=1}^{K}\xi_{ik}\phi_{k}(t)+\epsilon_i(t)\), where
\(\epsilon_i(t)\) indicates random noise introduced by this
approximation. This random noise is assumed to follow a zero-mean
Gaussian process, such that at each specific point t,
\(\epsilon_i(t) \sim N(0, \sigma^2)\).

In practice, this approximation could be implemented through an FPCA
procedure on \(\eta_i(t)\). In this case
\(\phi_{k}(t),\ k \in \{1,...,K\}\) are the first K eigenfunctions that
explain the most variation in \(\eta_i(t)\). These eigenfunctions, also
referred to as Principal Component (PC) functions, are orthogonal are
can be orthonormal with proper scaling. \(\xi_{ik}\) are
subject-specific scores (or loadings) on each PC function. According to
properties of FPCA method, these scores are mutually independent over
both subject (\(i\)) and eigenfunctions (\(k\)). That is, each
\(\xi_{ik}\) follows normal distribution \(N(0, \lambda_k)\), where
\(\lambda_k\) is the kth eigenvalues.

\textbf{Fast Generalized FPCA} Based on the problem set up above, we
propose the following algorithm for fast, generalizable implementation
of FPCA (fGFPCA) on the unobserved latent process \(\eta_i(t)\):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Bin the observed outcomes in to small, non-overlapping, equal length
  intervals. We hereafter index the bins by their midpoints \(s\). The
  number of observations from the same subject in bin \(s\) is \(n_s\),
  and \(t_{sj}\) refers to the jth observation point in bin \(s\).
\item
  Fit a local Generalized Mixed Model at every bin. Specifically, at bin
  \(s\), we fit \(g[E(Y_i(t_{sj}))] = \beta_0(s)+b_i(s)\),
  \(j\in\{1, ..., n_s\}\). From this series of models we can estimate
  the individual latent functions at every bin:
  \(\hat{\eta}_i(s) = \hat{\beta}_0(s)+\hat{b}_i(s)\).
\item
  Fit FPCA on the estimated latent functions:
  \(\hat{\eta}_i(s) = f_0(s)+\sum_{k=1}^K\xi_{ik}\phi_{k}(s)+\epsilon_i(s)\),
  and obtain estimates of basis functions
  \(\boldsymbol{\Phi} = \{\phi_1(s), ...,\phi_k(s)\}\), eigenvalues
  \(\hat{\lambda}_1...\hat{\lambda}_k\), mean function \(\hat{f}_0(s)\)
  and residual variance \(\hat{\sigma}^2\).
\end{enumerate}

\textbf{Dynamic prediction} Now assume we have a new observations with
partially observed outcome \(Y_{new}(t)\), \(t\leq t_m\). The maximum
observed point is \(t_m\), which is a point along the functional domain
and belongs to bin \(s_m\) on the binned grid from step 1 above. With
components extracted from the FPCA model in step 3, we will be able to
make predictions on the future, unobserved bins. Specifically, with the
partially observed binary function \(Y_{new}(t)\), we will be able to
calculate the maximum likelihood estimate (MLE) of the subject-specific
PC scores \(\hat{\xi}_{ik}\). Then the value of latent functions at
unobserved points can be estimated as
\(\hat{\eta_i}(s)=\hat{f}_0(s)+\sum_{k=1}^K\hat{\xi}_{ik}{\phi}_k(s)\)

Following the algorithm above, predictions of individual latent
functions are made on the binned grid based on partially observed tracks
on the un-binned grid. Since the bins are set up to be small in length,
the binned grid would still be dense. However, in situations where
predictions on the original, un-binned grid is needed, linear
interpolation of estimated latent function tracks turns out to be a
fast, convenient way with good performance for prediction at points
between bins.

\hypertarget{simulation}{%
\section{Simulation}\label{simulation}}

In this section, we illustrate the predictive performance and
computational efficiency of the proposed method through a simulation
study. We simulated 50 datasets, each with 500 subjects. For every
subject, we generate 1000 binary outcomes \(Y_i(t) \in \{0, 1\}\) across
functional domain \(t \in (0, 1000]\), where the distribution of outcome
is characterized by a continuous latent function. The data generation
mechanism can be expressed as follows:

\[\begin{aligned}
Y_i(t) & \sim Binomial(\frac{exp(\eta_i(t))}{1+exp(\eta_i(t))}) \\
\eta_i(t) &= f_0(t)+ \xi_{i1}\sqrt{2}sin(2\pi t)+\xi_{i2}\sqrt{2}cos(2\pi t)+\xi_{i3}\sqrt{2}sin(4\pi t)+\xi_{i4}\sqrt{2}cos(4\pi t)
\end{aligned}\]

In this simulation, we set \(f_0(t) = 0\). \(\xi_{ik}\) are mutually
independent normal random variables \(\xi_{ik}\sim N(0, \lambda_k)\).
Here we set the values of \(\lambda_k\) to be \(0.5^{k-1}\),
\(k \in (1,..., 4)\). In addition, for simplicity of presentation, we
generate data on a regular grid, which means observations points are
equally distributed across \((0, 1000]\) and are the same for all
subjects.

We use two metrics to evaluate the out-of-sample predictive performance:
integrated squared error (ISE) and
Area-Under-the-Receiver-Operator-Curve (AUC). ISE assesses the
prediction accuracy of latent continuous functions. It is evaluated on
the binned grid at midpoints of each unobserved bin. With observations
up to bin \(s_m\), ISE is defined as
\(\frac{1}{N}\sum_{i=1}^N\sum_{s>s_m} (\hat{\eta}_i(s)-\eta_i(s))^2\).
The second metric, AUC, focuses on evaluation of prediction of the
binary outcome. Since the binary outcomes are generated on the original,
un-binned grid, we evaluated AUC on this grid by estimating values of
latent functions between bins with linear interpolation.

As a reference method, we compare our method to Generalized Linear Mixed
Models using Adaptive Gaussian Quadrature (GLMMadaptive). This is one of
the fastest existing method developed for dynamic prediction of repeated
generalized outcomes. Just like many mixed models, this method is very
limited in terms of flexibility. For example, the model used for
prediction of our simulated datset would simply be an linear model with
one covariate indicating observation time:
\(g(E(Y_i)) = \beta_0+\beta_1t+b_{i0}+b_{i1}t\). While the flexibility
of this mixed model can be increased using spline functions, the
dimension of spline basis is also restricted by computational ability,
and is unfeasible to implement under the scale of our simulated data or
the complexity of our proposed method.

The average ISE and AUC across all simulation is presented in Table 1,
and Figure 1 presents the predicted latent function of four randomly
selected subjects from the first simulated dataset. The prediction is
made conditioning on different length of observed track (specifically,
with observations up to t = 200, 400, 600, 800 respectively), and
evaluation is made on equal-length time windows on the unobserved tracks
following the maximum observation time. As Table 1 reveals, the proposed
method (fGFPCA) outperforms GLMMadaptive under every scenario, which was
expected since the linear model structure of GLMM is too simple for the
cyclic data generation mechanism. In addition, fGFPCA also took less
computation time. For one simulated dataset, fGFPCA spent 4.94 minutes
on model fitting and out-of-sample prediction overall, while
GLMMadaptive took 5.11 minutes.

Because of the flexibility of our proposed model framework, the accuracy
of prediction at specific time points would improve as more observed
data is collected. This is revealed in Table 1 as ISE decreases and AUC
increases with maximum observed time (left-to-right), also in Figure 1
as predicted curves get closer to the true latent function with longer
observed track. However the same tendency is not observed with
GLMMadaptive models, as a result of restricted model flexibility. A
linear model that fits well to a specific part can fit very badly to
other parts along the underlying latent function tracks, especially when
the shapes of functions are wiggly or cyclic. Therefore, more
observations do not necessarily make the model more predictive.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/Figure-1.pdf}
\caption{Predicted track of four randomly selected subjects from one
simulated dataset. The grey solid line indicates the true latent
continuous function. The dashed lines indicated predicted latent
function tracks, and color indicates different observation time.}
\end{figure}

\begin{table}

\caption{\label{tab:format table}Predictive performance of fGFPCA and GLMM adaptive on the simulated datasets. ISE and AUC are average values across all 50 simulations.}
\centering
\begin{tabular}[t]{llccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{8}{c}{Maximum observed time} \\
\cmidrule(l{3pt}r{3pt}){2-9}
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{fGFPCA} & \multicolumn{4}{c}{GLMMadaptive} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & 200 & 400 & 600 & 800 & 200 & 400 & 600 & 800\\
\midrule
\addlinespace[0.3em]
\multicolumn{9}{l}{\textbf{Prediction time window}}\\
\addlinespace[0.3em]
\multicolumn{9}{l}{\textbf{ISE}}\\
\hspace{1em}\hspace{1em}(200, 400] & 15.29 &  &  &  & 38.20 &  &  & \\
\hspace{1em}\hspace{1em}(400, 600] & 18.68 & 8.53 &  &  & 28.59 & 26.46 &  & \\
\hspace{1em}\hspace{1em}(600, 800] & 22.53 & 5.51 & 1.65 &  & 31.32 & 27.97 & 27.40 & \\
\hspace{1em}\hspace{1em}(800, 1000] & 11.34 & 8.72 & 1.90 & 1.30 & 55.19 & 46.89 & 58.57 & 59.25\\
\addlinespace[0.3em]
\multicolumn{9}{l}{\textbf{AUC}}\\
\hspace{1em}\hspace{1em}(200, 400] & 0.74 &  &  &  & 0.59 &  &  & \\
\hspace{1em}\hspace{1em}(400, 600] & 0.66 & 0.73 &  &  & 0.52 & 0.59 &  & \\
\hspace{1em}\hspace{1em}(600, 800] & 0.71 & 0.79 & 0.80 &  & 0.67 & 0.70 & 0.69 & \\
\hspace{1em}\hspace{1em}(800, 1000] & 0.74 & 0.75 & 0.78 & 0.78 & 0.52 & 0.56 & 0.53 & 0.57\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{data-application}{%
\section{Data application}\label{data-application}}

In this section, we illustrate the performance of our proposed method
using data from a randomized trial on weight loss at University of
Colorado Anschutz Medical Campus
(\protect\hyperlink{ref-bothwell2022}{Bothwell et al. 2022};
\protect\hyperlink{ref-ostendorf2022}{Ostendorf et al. 2022}). The study
originally aimed to study the effect of dietary strategies on weight
loss, and participants were given a smart scale for daily at-home
weighing. This record was then scaled into a binary indicator of
participant adherence to the daily-weigh-in procedure, where 1 indicates
participants followed the requirement and 0 otherwise. In this section,
we will use this series of daily indicators as the binary functional
outcome, and use the proposed method to dynamically predict participant
adherence on future days given historical track.

The dataset includes 55 participants aged 22-56 at baseline. Adherence
to daily weigh-in is available along 400 consecutive days for all
participants. In other words, for each individual, 400 binary
measurements were taken on a regular grid along the functional domain
(number of days into the study). We will fit both fGFPCA and
GLMMadaptive models and compared their performance on this data. For
fGFPCA, every 10 observations are binned together, resulting in 40 bins
for final model fitting. Linear interpolation was used for extension
from the binned grid back to the un-binned grid. Since the latent
continuous function tracks are not known in this case, prediction
performance on evaluated with AUC alone.

Table 2 compares the predictive performance of both models. Similar to
the simulatio study, prediction is made conditioning on observations up
to the 100th, 200th and 300th day respectively, and evaluation is made
every 100 days following the maximum observation time. As is revealed,
when the observed time is short (up to 100th day), the two method
performed similarly when predicting recent days, however fGFPCA did
better in predicting further days along the trial (prediction time
window of \((300, 400]\)). As more observations coming in, fGFPCA sees
increasingly higher AUC compared to GLMMadaptive model. This could be a
result of the difference in model flexibility of the two methods. While
the linear relationship in underlying latent functions are not likely to
change immediately following the observed track, the whole track could
be much more complex. Therefore, while fGFPCA performs similarly to
GLMMadaptive right after the end of observation, it does a much better
job capturing the underlying pattern along the whole track. Figure 2
further illustrate such difference using predicted latent function
tracks of four randomly drawn subjects.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/Figure_appl-1.pdf}
\caption{Predicted track of four randomly selected subjects from the
daily weigh-in dataset. The grey solid line indicates the true latent
continuous function. The dashed lines indicated predicted latent
function tracks, and color indicates different observation time.}
\end{figure}

\begin{table}

\caption{\label{tab:unnamed-chunk-3}AUC of fGFPCA and GLMM adaptive on the daily weigh-in dataset}
\centering
\begin{tabular}[t]{llccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Maximum observed time} \\
\cmidrule(l{3pt}r{3pt}){2-7}
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{fGFPCA} & \multicolumn{3}{c}{GLMMadaptive} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
  & 100 & 200 & 300 & 100 & 200 & 300\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Prediction time window}}\\
\hspace{1em}(100, 200] & 0.75 &  &  & 0.76 &  & \\
\hspace{1em}(200, 300] & 0.77 & 0.82 &  & 0.76 & 0.80 & \\
\hspace{1em}(300, 400] & 0.76 & 0.81 & 0.84 & 0.74 & 0.78 & 0.82\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The simulation study and case study of daily weigh-in data above have
revealed the proposed method can achieve better predictive performance
with much less computational burden. Compared to existing methods based
on linear mixed model, fGFPCA can accommodate flexible correlation
structure between repeated measure, which is particularly useful when
outcome is generated from latent functions with wiggly or cyclic
pattern. As in the simulation setting in Section 3, the binary outcomes
are generated from a cyclic process. fGFPCA has much lower ISE and
higher AUC than GLMM in this case, and its prediction accuracy also
improves as more outcomes are collected, which is a desirable property
for dynamic prediction where the datasets would be periodically updated
with newly collected data. Moreover, fGFPCA also significantly reduce
time spent on model fitting and predict, which is beneficial when
repeated measures are dense or sample size is large. In fact, as far as
we are aware of, the proposed method is the only feasible method that
can handle the scale of the simulated datasets with compatible
flexibility.

While the proposed method achieved good accuracy of point prediction, it
is more complicated to develop a straightforward, interpretable metric
to evaluate the precision of prediction. The challenge arises from the
fact that fitting local GLMMs (Step 2) and FPCA (Step 3) can both
introduce uncertainty to the final prediction. Since they are
implemented consecutively, it is not clear how the uncertainty from both
procedures should be integrated into one single interval estimator. In
practice, one could consider developing interval predictions conditional
on the local GLMMs, quantifying uncertainty of individual score
estimates from FPCA model alone. Since the score \(\hat{\xi}_{ik}\) is
essentially a maximum likelihood estimate, it is natural to estimate its
variance based on likelihood theory with observed information
\({I}(\hat{\xi}_{ik})\). Therefore, the variation of prediction interval
is:

\[Var(\hat{\eta_i}(s)-\eta_i(s))=\boldsymbol{\Phi}(s)I(\hat{\boldsymbol{\xi}}_i)\boldsymbol{\Phi}^T(s)+\hat{\sigma_{\epsilon}}^2\]
Where \(\boldsymbol{\Phi}(s)=(\phi_1(s)...\phi_K(s))\) and
\(\hat{\boldsymbol{\xi}}_i=(\hat{\xi}_{i1}, ...,\hat{\xi}_{iK})\).

The hyperparameters introduced by the binning procedure in Step 1 can
also affect the final predictive performance, thus further
investigations are needed to quantify such effects. Predictive accuracy
and computational speed can change with bin width, number of
observations in each bin, whether the bins overlap with each other, also
whether the bins are equally spaced along the functional grid. Moreover,
if the bins are very narrow with only a few observations,
unidentifiability issues can happen when fitting local mixed models.
While this may not necessarily affect point predictions, it can cause
numeric failure to estimate standard error and prediction intervals. In
practice, we expect this to happen when the measurements are sparse or
sample size is small. Under such scenarios, one may choose traditional
mixed model framework instead to get valid estimates at the expense of
computational time.

It should also be noted that the binning procedure in Step 1 changes the
density of repeated measures. We essentially assume that latent
functions are smooth in small intervals, and treat adjacent observations
as if they are generated from the same time point. The final prediction
of latent continuous function tracks are made on the binned grid,
integrating information from several adjacent observations in the same
bin. In Section 3 and 4, this sparser grid was transformed to the
original, un-binned grid using linear interpolation of the predicted
latent function. However, this is by no means the only method to be
considered. Such transformation can also be made by projection to a
higher-dimensional space, similar as general additive model with a large
number of basis functions. We suspect that this method would work better
than linear interpolation when the original grid is sparse or irregular,
but take longer time for computation.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-bothwell2022}{}}%
Bothwell, S., Kaizer, A., Catenacci, V., and Wrobel, J. (2022),
{``Pattern-based clustering of daily weigh-in trajectories using dynamic
time warping,''} \emph{Biometrics}.
\url{https://doi.org/10.1111/biom.13773}.

\leavevmode\vadjust pre{\hypertarget{ref-chen2013}{}}%
Chen, H., Wang, Y., Paik, M. cho, and Choi, H. A. (2013), {``A marginal
approach to reduced-rank penalized spline smoothing with application to
multilevel functional data,''} \emph{J Am Stat Assoc.}, 108, 1216--1229.
\url{https://doi.org/10.1080/01621459.2013.826134}.

\leavevmode\vadjust pre{\hypertarget{ref-chiou2012}{}}%
Chiou, J.-M. (2012), {``Dynamical functional prediction and
classification, with application to traffic flow prediction,''}
\emph{The Annals of Applied Statistics}, Institute of Mathematical
Statistics, 6, 1588--1614. \url{https://doi.org/10.1214/12-AOAS595}.

\leavevmode\vadjust pre{\hypertarget{ref-delaigo2016}{}}%
Delaigle, A., and Hall, P. (2016), {``Approximating fragmented
functional data by segments of markov chains,''} \emph{Biometrika}, 103,
779--799. \url{https://doi.org/10.1093/biomet/asw040}.

\leavevmode\vadjust pre{\hypertarget{ref-gertheiss2016}{}}%
Gertheiss, J., Goldsmith, J., and Staicu, A. (2016), {``A note on
modeling sparse exponential-family functional response curves,''}
\emph{Comput Stat Data Anal}, 105, 46--52.
\url{https://doi.org/10.1016/j.csda.2016.07.010}.

\leavevmode\vadjust pre{\hypertarget{ref-goldberg2014}{}}%
Goldberg, Y., Ritov, Y., and Mandelbaum, A. (2014), {``Predicting the
continuation of a function with applications to call center data,''}
\emph{Journal of Statistical Planning and Inference}, 147, 53--65.
https://doi.org/\url{https://doi.org/10.1016/j.jspi.2013.11.006}.

\leavevmode\vadjust pre{\hypertarget{ref-goldsmith2015}{}}%
Goldsmith, J., Zipunnikov, V., and Schrack, J. (2015), {``Generalized
multilevel function-on-scalar regression and principal component
analysis,''} \emph{Biometrics}, 71, 344--53.
\url{https://doi.org/10.1111/biom.12278}.

\leavevmode\vadjust pre{\hypertarget{ref-hall2018}{}}%
Hall, P., Müller, H.-G., and Yao, F. (2008), {``Modelling sparse
generalized longitudinal observations with latent gaussian processes,''}
\emph{Journal of the Royal Statistical Society: Series B (Statistical
Methodology)}, 70, 703--723.
https://doi.org/\url{https://doi.org/10.1111/j.1467-9868.2008.00656.x}.

\leavevmode\vadjust pre{\hypertarget{ref-kraus2015}{}}%
Kraus, D. (2015),
{``\href{http://www.jstor.org/stable/24775309}{Components and completion
of partially observed functional data},''} \emph{Journal of the Royal
Statistical Society. Series B (Statistical Methodology)}, {[}Royal
Statistical Society, Wiley{]}, 77, 777--801.

\leavevmode\vadjust pre{\hypertarget{ref-Laird1982}{}}%
Laird, N. M., and Ware, J. H. (1982),
{``\href{http://www.jstor.org/stable/2529876}{Random-effects models for
longitudinal data},''} \emph{Biometrics}, {[}Wiley, International
Biometric Society{]}, 38, 963--974.

\leavevmode\vadjust pre{\hypertarget{ref-leroux2022}{}}%
Leroux, A., Crainiceanu, C. M., and Wrobel, J. (n.d.). {``Fast
generalized functional principal component analysis.''}

\leavevmode\vadjust pre{\hypertarget{ref-leroux2016}{}}%
Leroux, A., Xiao, L., Crainiceanu, C., and Checkley, W. (2018),
{``Dynamic prediction in functional concurrent regression with an
application to child growth,''} \emph{Statistics in medicine}, 37,
1376--1388.

\leavevmode\vadjust pre{\hypertarget{ref-liang1986}{}}%
LIANG, K.-Y., and ZEGER, S. L. (1986), {``Longitudinal data analysis
using generalized linear models,''} \emph{Biometrika}, 73, 13--22.
\url{https://doi.org/10.1093/biomet/73.1.13}.

\leavevmode\vadjust pre{\hypertarget{ref-linde2019}{}}%
Linde, van der (2009), {``A bayesian latent variable approach to
functional principal components analysi with binary and count data,''}
\emph{A StA Adv Stat Anal}, 307--333.
\url{https://doi.org/10.1007/s10182-009-0113-6}.

\leavevmode\vadjust pre{\hypertarget{ref-lindstrom1990}{}}%
Lindstrom, M. J., and Bates, D. M. (1990),
{``\href{http://www.jstor.org/stable/2532087}{Nonlinear mixed effects
models for repeated measures data},''} \emph{Biometrics}, {[}Wiley,
International Biometric Society{]}, 46, 673--687.

\leavevmode\vadjust pre{\hypertarget{ref-davidian2003}{}}%
{``\href{http://www.jstor.org/stable/1400665}{Nonlinear models for
repeated measurement data: An overview and update}''} (2003),
{[}International Biometric Society, Springer{]}, 8, 387--419.

\leavevmode\vadjust pre{\hypertarget{ref-ostendorf2022}{}}%
Ostendorf, D. M., Caldwell, A. E., Zaman, A., Pan, Z., Bing, K.,
Wayland, L. T., Creasy, S. A., Bessesen, D. H., MacLean, P., Melanson,
E. L., and Catenacci, V. A. (2022), {``Comparison of weight loss induced
by daily caloric restriction versus intermittent fasting (DRIFT) in
individuals with obesity: Study protocol for a 52-week randomized
clinical trial,''} \emph{Trials}, 23, 718.
\url{https://doi.org/10.1186/s13063-022-06523-2.}

\leavevmode\vadjust pre{\hypertarget{ref-GLMMadaptive}{}}%
Rizopoulos, D. (2022),
\emph{\href{https://CRAN.R-project.org/package=GLMMadaptive}{GLMMadaptive:
Generalized linear mixed models using adaptive gaussian quadrature}}.

\leavevmode\vadjust pre{\hypertarget{ref-Scheipl2014}{}}%
Scheipl, F., Staicu, A.-M., and Greven, S. (2014), {``Functional
additive mixed models,''} \emph{J Comput Graph Stat}, 24, 447--501.
\url{https://doi.org/10.1080/10618600.2014.901914}.

\leavevmode\vadjust pre{\hypertarget{ref-shang2017}{}}%
Shang, H. L. (2017), {``Functional time series forecasting with dynamic
updating: An application to intraday particulate matter
concentration,''} \emph{Econometrics and Statistics}, 1, 184--200.
https://doi.org/\url{https://doi.org/10.1016/j.ecosta.2016.08.004}.

\leavevmode\vadjust pre{\hypertarget{ref-suresh2017}{}}%
Suresh, K., Taylor, J. M. G., Spratt, D. E., Daignault, S., and
Tsodikov, A. (2017), {``Comparison of joint modeling and landmarking for
dynamic prediction under an illness-death model,''} \emph{Biom J}, 59,
1277--1300. \url{https://doi.org/10.1002/bimj.201600235}.

\leavevmode\vadjust pre{\hypertarget{ref-wrobel2019}{}}%
Wrobel, J., Zipunnikov, V., Schrack, J., and Goldsmith, J. (2019),
{``Registration for exponential family functional data,''}
\emph{Biometrics}, 75, 48--57.
https://doi.org/\url{https://doi.org/10.1111/biom.12963}.

\end{CSLReferences}

\end{document}