### 2023-09-27 Docs/ProgressReport.Rmd
try splines interpolation of mean and eigenfunctions

---
title: "Progress Report"
author: "Ying Jin"
date: "2023-09-15"
output: 
  html_document:
    self_contained: no
    number_sections: true
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    font: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(516)

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(mvtnorm)
library(mgcv)
library(splines)
theme_set(theme_minimal())

df <- read_rds(here("Data/nhanes_bi.rds"))
# load(here("Data/ApplOutput_fGFPCA.RData"))

```

# NHANES data

## Data overview

- 8763 subjects, 1440 measures each
- no missingness

```{r}
N <- length(unique(df$SEQN)) # sample size 8763
J <- max(df$sind) # 1440 measures per subject
```

```{r}
rand_id <- sample(unique(df$SEQN), size = 4) # "toy" sample

df %>% 
  filter(SEQN %in% rand_id) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Z), size = 0.5)+
  facet_wrap(~SEQN)+
  labs(x="Time", y = "Activity", title = "A brief overview of the outcome")
```

## fGFPCA model fitting

```{r}
# code
source(here("Code/GLMM-FPCA.R")) 
# use pred_latent function to estimate latent function 
source(here("Code/OutSampMLE.R"))
# source(here("Code/OutsampBayes.R"))
```

### Bin Data

- Bin every 10 consecutive observations

```{r, message=FALSE}
# bin data
bin_w <- 10 # bin width
n_bin <- J/bin_w # number of bins
brks <- seq(0, J, by = bin_w) # cutoff points
mid <- (brks+bin_w/2)[1:n_bin] # mid points

df$bin <- cut(df$sind, breaks = brks, include.lowest = T, labels = mid)
df$bin <- as.numeric(as.character(df$bin))
# head(df)

df %>% 
  filter(SEQN %in% rand_id) %>%
  group_by(SEQN, bin) %>%
  summarise(num = sum(Z)) %>%
  ggplot()+
  geom_point(aes(x=bin, y=num), size = 0.5)+
  facet_wrap(~SEQN)+
  labs(x="Time", y = "Activity", title = "Number of active nimutes within each bin")

df <- df %>% rename(id = SEQN, Y=Z)
```
### Data split

```{r}
id_vec <- unique(df$id)
train_id <- sample(id_vec, size = 0.8*N, replace = FALSE) # 7010 subjects for training
test_id <- setdiff(id_vec, train_id) # 1753 subjects for testing
```

- 80% (7017) subjects for model fitting, 20% (1753) subjects for out-of-sample prediction

### Local GLMM

- Used glmer with PIRLS (nAGQ=0) for local GLMMs step
- Fit on the training set

```{r, cache=TRUE}
# fit model on the training set
df_train <- df %>% filter(id %in% train_id)
train_bin_lst <- split(df_train, f = df_train$bin)

# local GLMM and estimate latent function
# use PIRLS (nAGQ=0) to avoid near-unidentifiability issues 
t1=Sys.time()
df_est_latent <- lapply(train_bin_lst, function(x){pred_latent(x, n_node = 0)}) 
t2= Sys.time()
t_local_glmm <- t2-t1 
```

```{r}
df_est_latent <- bind_rows(df_est_latent) %>% 
  select(-sind, -Y) %>% distinct(.) 
# example estimated latent function
df %>% 
  filter(id %in% rand_id) %>%
  left_join(df_est_latent, by = c("id", "bin")) %>%
  mutate(eta_hat = exp(eta_hat)/(1+exp(eta_hat))) %>%
  ggplot()+
  geom_line(aes(x=sind, y=eta_hat, group = id))+
  geom_point(aes(x=sind, y = Y, group = id), size = 0.5)+
  facet_wrap(~id, scales = "free")+
  labs(x = "Time", y = "Estimated latent function (probablity scale)")
```

```{r, message=FALSE}

df %>% 
  filter(id %in% sample(train_id, size = 4)) %>%
  group_by(id, bin) %>%
  summarise(num = sum(Y)) %>%
  left_join(df_est_latent, by = c("id", "bin")) %>%
  ggplot()+
  geom_point(aes(x=bin, y=num), size = 0.5)+
  geom_line(aes(x=bin, y=eta_hat, group = id))+
  facet_wrap(~id)+
  labs(x="Time", y = "Estimated latent function (original scale")
```

### FPCA

```{r}
mat_est_unique <- matrix(df_est_latent$eta_hat,
                         nrow=length(train_id), 
                         ncol=n_bin, byrow = F) # row index subject, column binned time
#dim(mat_est_unique)

t1 <- Sys.time()
fpca_mod <- fpca.face(mat_est_unique, argvals = mid, var=T)
t2 <- Sys.time()
t_fpca <- t2-t1 # 3.21 seconds to fit fPCA model

# dim(fpca_mod$efunctions) # 27 eigenfunctions total
# fpca_mod$evalues
K <- 4
```

```{r}
plot(mid, fpca_mod$mu, type = "l", xlab = "bin", ylab = "mean")
```

```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))

plot(mid, fpca_mod$efunctions[, 1], type="l", xlab="bin", ylab="PC1")
plot(mid, fpca_mod$efunctions[, 2], type="l", xlab="bin", ylab="PC2")
plot(mid, fpca_mod$efunctions[, 3], type="l", xlab="bin", ylab="PC3")
plot(mid, fpca_mod$efunctions[, 4], type="l", xlab="bin", ylab="PC4")
```

```{r}
# plot correlation matrix
heatmap(cov2cor(fpca_mod$VarMats[[1]]), Rowv = NA, Colv = NA, main = "Correlation")
```

### Debias (step 4)

According to the fGFPCA paper, in this step we do two things:

1. Interpolate the eigenfunctions with B-spline basis back to the original grid
2. Debias the eigenvalues with GLMM (mgcv::bam, mgcm::gamm or gamm4:gamm) to be used for Laplace approximation
- we do not need to debias individual scores because the scores for training sample will not be used for out-of-sample prediction

First I'd like to try to fit the debias GLMM model on the full training set with un-interpolated eigenfunctions. The model failed because of vector memory exhaustion. Shall I: 
- Decrease the density of grid (increase bin width)
- Increase memory (re-set to 100GB). But still took a lot of time (more than one hour). The model actually did not finish fitting.

The bias was caused by the misspecification of latent process (assume constant effect over smooth function). And what needs re-evaluation are eigenvalues. Basis functions and mean functions are not biased.

```{r, eval=FALSE}
df_phi <- data.frame(bin=mid, fpca_mod$efunctions[, 1:4])
colnames(df_phi) <- c("bin", paste0("phi", 1:4))

df_train <- df_train %>% left_join(df_phi, by = "bin")
df_train$id <- as.factor(df_train$id)

# usethis::edit_r_environ()
t1 <- Sys.time()
debias_glmm <- bam(Y ~ s(bin, bs="cr", k=10)+
                     s(id, by=phi1, bs="re")+
                     s(id, by=phi2, bs="re")+
                     s(id, by=phi3, bs="re")+
                     s(id, by=phi4, bs="re"), 
                   family = binomial, data=df_train, method = "fREML")
t2 <- Sys.time()
t_debias <- t2=t1

```

Then I would like to interpolate eigenfunctions using B-spline basis functions:

- Would you interpolate mean and eigenfunctions separately (in my case, five interpolations)? 
- I did not find the reeval_efunctions function

```{r, fig.height=5, fig.width=5}
# mean
interp_mean <- interpSpline(mid, fpca_mod$mu)
full_mean <- predict(interp_mean, x=1:J)
plot(mid, fpca_mod$mu, ylab = "mean", pch = 20)
lines(full_mean$x, full_mean$y, col = "blue")

# PCs
interp_phi1 <- interpSpline(mid, fpca_mod$efunctions[, 1])
full_phi1 <- predict(interp_phi1, x=1:J)
plot(mid, fpca_mod$efunctions[, 1], ylab = "pc1", pch=20)
lines(full_phi1$x, full_phi1$y, col = "blue")

interp_phi2 <- interpSpline(mid, fpca_mod$efunctions[, 2])
full_phi2 <- predict(interp_phi2, x=1:J)
plot(mid, fpca_mod$efunctions[, 2], ylab = "pc2", pch=20)
lines(full_phi2$x, full_phi2$y, col = "blue")

interp_phi3 <- interpSpline(mid, fpca_mod$efunctions[, 3])
full_phi3 <- predict(interp_phi3, x=1:J)
plot(mid, fpca_mod$efunctions[, 3], ylab = "pc3", pch=20)
lines(full_phi3$x, full_phi3$y, col = "blue")

interp_phi4 <- interpSpline(mid, fpca_mod$efunctions[, 4])
full_phi4 <- predict(interp_phi4, x=1:J)
plot(mid, fpca_mod$efunctions[, 4], ylab = "pc4", pch=20)
lines(full_phi4$x, full_phi4$y, col = "blue")
```

## Out-of-sample prediction

follow up on the last report, I'll add some details here.

Let's set:

- $s$: bin index
- $t_{m_s}$: midpoint time of bin s
- $\mathscr{T}_s$: all observation points in bin s
- $Y_i^s$ all observations in bin s from subject i. $Y_i^S := \{Y_i(t_j), t_j \in \mathscr{T}_s\}$

Let's assume we have a new subject $u$ with maximum observation time $T_u$. Then the log-likelihood of this new subject would be:

$$
l_u=\sum_{t_{m_s}<T_u}log(h(Y_u^s))+\hat{\eta}_u(t_{m_s})T(Y_u^s)-log(A[\hat{\eta}_u(t_{m_s})])
$$

where $\hat{\eta}_u(t_{m_s}) = \hat{\mu}_0(t_{m_s})+\sum_{k=1}^K \xi_{uk}\hat{\phi}(t_{m_s})$.

From FPCA model (or the debias step to be added), we obtain $\hat{\mu}_0$, $\hat{\phi}_k$, as well as the variance estimates of $\xi_k$: $\hat{\lambda}_k$, and variance of residual process $\hat{\sigma^2}$.

With all these estimates, we wanna use Laplace approximation to find the $\xi_{uk}$ that maximizes $l_u$.

I am using a package called **LaplacesDemon**, following this example https://cran.r-project.org/web/packages/LaplacesDemon/vignettes/LaplacesDemonTutorial.pdf

```{r, class.source = "fold-show"}
library(LaplacesDemon)
```

Now I am going to use one subject as an example to go over the approximation procedure:

First we write out the model:

- Prior distribution: $\xi_{uk} \sim N(0, \hat{\lambda}_k)$
- Posterior distribution:

$$\begin{aligned}
l(\mathbf{Y_u}|\mathbf{\xi}_u) &= \sum_{t_{m_s}<T_u}l(Y_u^s|\mathbf{\xi}_u) \\

Y_u^s|\mathbf{\xi}_u & \sim Binomial(n_s, p_s) \\

g^{-1}(p_s) = \hat{\eta}(t_{m_s}) &= \hat{\mu}_0(t_{m_s})+\sum_{k=1}^K \xi_{uk}\hat{\phi}(t_{m_s})
\end{aligned}$$

And $n_s$ is the number of observations in bin s.

Bayes theroem:

$$
l(\mathbf{\xi}_u|\mathbf{Y}_u) \propto l(\mathbf{Y}_u|\mathbf{\xi}_u)+l(\mathbf{\xi}_u)
$$

```{r, class.source = "fold-show"}
## model
Model <- function(parm, Data){
  xi <- parm[Data$pos.xi]

  # log-prior
  xi.prior <- dmvnorm(xi, mean = rep(0, Data$J), sigma=Data$tao, log = TRUE)

  # log-posterior likelihood
  eta <- Data$f0+Data$X %*% xi
  p <- exp(eta)/(1+exp(eta))
  LL <- sum(dbinom(x=Data$y, size = Data$n, prob=p, log = TRUE)) # log likelihood of Y|xi
  LP <- LL+sum(xi.prior) # unnormalized joint log likelihood of (Y, xi)

  # output
  Modelout <- list(LP=LP, Dev=-2*LL, Monitor=LP,
                   yhat=Data$y, parm=parm)
  return(Modelout)
}
```

- Then we feed in data for one observation

```{r, class.source = "fold-show"}
rand_id <- test_id[15]
df_new <- df %>% filter(id == rand_id & sind <= 540) # observation until 9am

# put data into correct format
ns <- as.vector(table(df_new$bin)) # number of observations
hs <- df_new %>% group_by(bin) %>% summarize_at("Y", sum) %>% select(Y) %>% unlist()# number of success
nf <- ns-hs # number of failure
max_bin <- length(unique(df_new$bin)) # assume new skipped bins  
df_new2 <- data.frame(bin = unique(df_new$bin), ns, hs, nf)

# extract estimates from FPCA model to be used in prior distribution
tao <- diag(fpca_mod$evalues[1:K]) # variance of xi

# extract estimates from FPCA model to be used in posterior distribution
f0 <- fpca_mod$mu[1:max_bin]
N <- nrow(df_new2) # "sample size" which is in fact number of observed bins in this case for a new subject
n <- df_new2$ns # number of experiements at each bin
y <- df_new2$hs # outcome, number of 1
X <- fpca_mod$efunctions[1:max_bin, 1:K] # eigen/PC functions
J <- K # number of parameters/scores
mon.names <- "LP"

# parameter names
name_lst <- as.list(rep(0, J))
names(name_lst) <- paste("xi", 1:J, sep = "")
parm.names <- as.parm.names(name_lst) # names of parameters to estimate
pos.xi <- grep("xi", parm.names)

# data
MyData <- list(J=J, X=X, mon.names=mon.names,
                 parm.names=parm.names, pos.xi=pos.xi, y=y, n=n, tao=tao, f0=f0)
```

```{r, class.source = "fold-show"}
# fit laplace approximation
# initial values generated by GIV function in the same package
Fit <- LaplaceApproximation(Model, Data=MyData)
```

- Summarizing output

```{r, class.source = "fold-show", fig.height=8, fig.width=8}
# summarizing output
print(Fit)
plot(Fit, MyData)
```

```{r}
# prediction
eta_pred_out <- fpca_mod$mu+fpca_mod$efunctions[, 1:K]%*%Fit$Summary1[, "Mode"]

data.frame(bin = mid, eta_pred = eta_pred_out) %>% 
  right_join(df %>% filter(id==rand_id) %>%  select(Y, sind, bin), by = "bin") %>%
  ggplot()+
  geom_point(aes(x=sind, y = Y))+
  geom_line(aes(x=sind, y = exp(eta_pred)/(1+exp(eta_pred))))+
  labs(title = rand_id)

```

# Full data application output

## fGFPCA

- Still need to project eigenfunctions to the full grid

```{r}
# results generated from DataApplNHANES.R
load(here("Data/ApplOutput_fGFPCA.RData"))
```

```{r message=FALSE, warning=FALSE}
df_test_full <- df %>% 
  filter(id %in% test_id) %>%
  filter(!id %in% skip_id) %>%
  left_join(df_test %>% select(id, bin, pred_t540, pred_t780, pred_t1020), by = c("id", "bin"))

rand_id2 <- sample(unique(df_test$id), 4)
# without interpolation, assume constant latent function value in each bin
df_test_full%>%
  filter(id %in% rand_id2) %>%
  mutate_at(vars(pred_t540, pred_t780, pred_t1020), function(x)exp(x)/(1+exp(x))) %>%
  ggplot()+
    geom_line(aes(x=bin, y = pred_t540, col = "9am"))+
    geom_line(aes(x=bin, y = pred_t780, col = "1pm"))+
    geom_line(aes(x=bin, y = pred_t1020, col = "5pm"))+
    geom_point(aes(x=bin, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "Estimated latent function (probablity scale)",
         title = "Constant latent function within bin")
```

### Numeric problems

- Three subjects had failed approximation, all happened with the shortest observed track
- Does it have to do with still the length of observed track? Or the arbitrary choices in the binning procedure (bin width). 
- Per last meeting, adding some details on laplace approximation

```{r, fig.height=3, fig.width=9}
df %>% filter(id %in% skip_id) %>% 
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  facet_wrap(~id)+
  geom_vline(xintercept = c(540))
```

<!-- Calculate AUC -->

```{r}
# break by prediction window
lst_auc <- df_test_full %>% 
  mutate(window = cut(sind, breaks = c(0, 540, 780, 1020, 1260, 1440), 
                      labels = c("0-9am", "9am-1pm", "1pm-5pm", "5pm-9pm","9pm-12pm"), 
                      include.lowest = T))
lst_auc <- split(lst_auc, f=lst_auc$window)
```

```{r}
# no interpolation
# up to 9am
auc_t540 <- lst_auc[2:4] %>% 
  lapply(function(x)performance(prediction(x$pred_t540, x$Y), measure = "auc"))
auc_t540 <- lapply(auc_t540, function(x){x@y.values[[1]]}) %>% unlist()

# up to 1pm
auc_t780 <- lst_auc[3:4] %>% 
  lapply(function(x)performance(prediction(x$pred_t780, x$Y), measure = "auc"))
auc_t780 <- lapply(auc_t780, function(x){x@y.values[[1]]}) %>% unlist()

# up to 5pm
auc_t1020 <- lst_auc[4] %>% 
  lapply(function(x)performance(prediction(x$pred_t1020, x$Y), measure = "auc"))
auc_t1020 <- lapply(auc_t1020, function(x){x@y.values[[1]]}) %>% unlist()
```

## GLMMadaptvie

- when fitting model with **random intercept + slope** on the same full training set (even after scaling covariates): Error: vector memory exhausted (limit reached?).
- when fitting model with **only random intercept** on the same full training set: Error regarding a large coefficient value. One fix is to re-scale covariates. So I re-scaled minute index by dividing them by J, so the range is within (0, 1], and tried again. This fitting procedure took 20.72 minutes to finish

```{r}
load(here("Data/ApplOutput_GLMMadaptive.RData"))
```

- `r round(t_est_adglmm, 2)` minutes on model fitting and `r round(t_pred_adglmm, 2)` minutes on out-of-sample prediction.

- Below is the model fit on the training set

```{r}
summary(adglmm_mod)
```

```{r}
df_test_full2 <- df_test_full %>%
  filter(id %in% rand_id2)%>%
  select(id, sind, Y)

df_test_full2$pred_t540 <- df_test_full2$pred_t780 <- df_test_full2$pred_t1020 <- NA 
for(i in rand_id2){
  df_test_full2[df_test_full2$id==i & df_test_full2$sind>540, "pred_t540"]<- adglmm_pred_t540$newdata2 %>% filter(id ==i) %>% select(pred)

  df_test_full2[df_test_full2$id==i & df_test_full2$sind>780, "pred_t780"]<- adglmm_pred_t780$newdata2 %>% filter(id ==i) %>% select(pred)

  df_test_full2[df_test_full2$id==i & df_test_full2$sind>1020, "pred_t1020"]<- adglmm_pred_t1020$newdata2 %>% filter(id ==i) %>% select(pred)
}

# colSums(is.na(df_test_full2))
```

- Below is an example of predicted probablity on four random subjects

```{r, warning=FALSE}
df_test_full2 %>% 
  mutate_at(vars(pred_t540, pred_t780, pred_t1020), 
            function(x)exp(x)/(1+exp(x))) %>%
  ggplot()+
    geom_line(aes(x=sind, y = pred_t540, col = "9am"))+
    geom_line(aes(x=sind, y = pred_t780, col = "1pm"))+
    geom_line(aes(x=sind, y = pred_t1020, col = "5pm"))+
    geom_point(aes(x=sind, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "Estimated latent function (probablity scale)")
```

<!-- Calculate AUC  -->

```{r}
# up to 540
J <- 1440
auc_t540_adglmm <- adglmm_pred_t540$newdata2 %>%
  filter(!id %in% skip_id) %>%
  mutate(sind=sind*J) %>%
  mutate(window = cut(sind, breaks = c(540, 780, 1020, 1260, 1440), 
                      labels = c("9am-1pm", "1pm-5pm", "5pm-9pm","9pm-12pm"), 
                      include.lowest = T))
auc_t540_adglmm <- split(auc_t540_adglmm, f=auc_t540_adglmm$window)
auc_t540_adglmm <- auc_t540_adglmm[1:3] %>% 
  lapply(function(x)performance(prediction(x$pred, x$Y), measure = "auc"))
auc_t540_adglmm <- lapply(auc_t540_adglmm, function(x){x@y.values[[1]]}) %>% unlist()

# up to 780
auc_t780_adglmm <- adglmm_pred_t780$newdata2 %>%
  filter(!id %in% skip_id) %>%
  mutate(sind=sind*J) %>%
  mutate(window = cut(sind, breaks = c(780, 1020, 1260, 1440), 
                      labels = c("1pm-5pm", "5pm-9pm","9pm-12pm"), 
                      include.lowest = T))
auc_t780_adglmm <- split(auc_t780_adglmm, f=auc_t780_adglmm$window)
auc_t780_adglmm <- auc_t780_adglmm[1:2] %>% 
  lapply(function(x)performance(prediction(x$pred, x$Y), measure = "auc"))
auc_t780_adglmm <- lapply(auc_t780_adglmm, function(x){x@y.values[[1]]}) %>% unlist()

# up to 1020
auc_t1020_adglmm <- adglmm_pred_t1020$newdata2 %>%
  filter(!id %in% skip_id) %>%
  mutate(sind=sind*J) %>%
  mutate(window = cut(sind, breaks = c(1020, 1260, 1440), 
                      labels = c("5pm-9pm","9pm-12pm"), 
                      include.lowest = T))
auc_t1020_adglmm <- split(auc_t1020_adglmm, f=auc_t1020_adglmm$window)
auc_t1020_adglmm <- auc_t1020_adglmm[1] %>% 
  lapply(function(x)performance(prediction(x$pred, x$Y), measure = "auc"))
auc_t1020_adglmm <- lapply(auc_t1020_adglmm, function(x){x@y.values[[1]]}) %>% unlist()

```

## Compare fGFPCA and GLMMadaptive

- To use the same sample for comparison, I'd exclude the subjects with failed Laplace approximation from the test set for both methods

```{r}
options(knitr.kable.NA = '')
tb_auc <- data.frame(auc_t540, 
           c(NA, auc_t780), 
           c(NA, NA, auc_t1020),

           auc_t540_adglmm,
           c(NA, auc_t780_adglmm), 
           c(NA, NA, auc_t1020_adglmm))

colnames(tb_auc) <- rep(c("9am", "1pm", "5pm"), 2)
```

```{r}
# time
t_fit <- c(t_local_glmm+t_fpca, t_est_adglmm)
units(t_fit) <- "mins"

t_test <- c(t_pred, t_pred_adglmm)
# units(t_pred)

t_total <- t_fit+t_test

idx1 <- idx2 <- idx3 <- c(1, 3, 3)
names(idx1) <- c("Total time", as.character(round(t_total, 2)))
names(idx2) <- c("Time on prediction", as.character(round(t_test, 2)))
names(idx3) <- c("Time on model fitting", as.character(round(t_fit, 2)))

t_fit <- as.character(round(t_fit, 2))
t_test <- as.character(round(t_test, 2))
t_total <- as.character(round(t_total, 2))

```

```{r}
tb_auc %>%
  kable(digit = 4, table.attr = "style = \"color: black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(header=idx1) %>%
  add_header_above(header=idx2) %>%
  add_header_above(header=idx3) %>%
  add_header_above(c(" "=1, "fGFPCA"=3, "GLMMadaptive"=3))
```

- fGFPCA does not always have higher AUC, but definitely most of the time. 
- In fact there are two scenarios where GLMMadaptive outperforms fGFPCA slightly: 1) given 9am to predict 1-5pm; and 2) given 1pm and predict 5-9pm. Both of them have a 4-hour interval apart from the maximum observation time and and prediction window. Could it have anything to do the cyclic trend in the data? Is 4h the length of a cycle? I tried to look at it but really couldn't see anything...there are too many points lumped together!

```{r}
df %>% filter(id %in% rand_id2) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  geom_vline(xintercept = c(540, 780, 1020, 1260))+
  scale_x_continuous(breaks = c(540, 780, 1020, 1260),
                     labels = c("9am", "1pm", "5pm", "9pm"))+
  facet_wrap(~id)
```

## Discussion

- What would be the reference method to compare our performance to? Original GLMMadaptvie perhaps is feasible on the entire dataset. But what should we use if we wanna do sub-sample comparison? Should we even do that for this data application (or simulation alone)? 
- There are still three participants with numeric problem! Shall we extend the observation track even more? I suspect that may cause still the same issue just on different subjects. 
- Is interpolation really necessary? It does not improve performance very much but consumes time. Also, should I try more sophisticated methods for grid extension? 
- Should I calculate subject-wise AUC and average them? But some subjects are all zero within the prediction window, making AUC for this single subjects impossible to calculate...