### 2024-03-26 Docs/NhanesApplExample.Rmd
repeated simulation use NHANES estimates

---
title: "Reverse eigenvalue problem"
author: "Ying Jin"
output: 
  html_document:
    self_contained: yes
    number_sections: true
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    font: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(308)

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(mvtnorm)
library(mgcv)
library(splines)
library(LaplacesDemon)
library(arsenal)
library(RColorBrewer)
theme_set(theme_minimal())
```

```{r load_data}
# load data
df <- read_rds(here("Data/nhanes_bi.rds"))
df <- df %>% rename(id=SEQN, Y=Z)

N <- length(unique(df$id))
J <- length(unique(df$sind))
K <- 4 # number of eigenfunctions to use
```

```{r load_code}
source(here("Code/GLMM-FPCA.R")) 
source(here("Code/OutsampBayes.R"))
```

It seems that the reversed eigenvalue problem has to do with 1) the choice of spline basis and 2) sample size.

# Truncated track

I'll take a subset of 500 subject and try different spline basis. In addition, I truncated the dataset to keep records during the daytime (8am-6pm).

```{r data_split}
# for sanity checks, use a subset of 500 subjects
Ntry <- 500
train_id <- sample(unique(df$id), size = Ntry)
train_df <- df %>% filter(id %in% train_id)

# only the day time? 
# 8am - 6pm (sind: 480 - 1080)
train_df <- train_df %>% filter(sind > 480 & sind <= 1080)
train_df$sind <- train_df$sind-480
J <- max(train_df$sind)
```

```{r}
# Step 1: Bin every 10 observations
bin_w <- 10 # bin width
n_bin <- J/bin_w # number of bins
brks <- seq(0, J, by = bin_w) # cutoff points
mid <- (brks+bin_w/2)[1:n_bin] # mid points

train_df$bin <- cut(train_df$sind, breaks = brks, include.lowest = T, labels = mid)
train_df$bin <- as.numeric(as.character(train_df$bin))

# Step 2:  Local GLMM
train_bin_lst <- split(train_df, f = train_df$bin)
df_est_latent <- lapply(train_bin_lst, function(x){pred_latent(x, n_node = 0)}) 
df_est_latent <- bind_rows(df_est_latent)

# FPCA
uni_eta_hat <- df_est_latent %>% filter(bin==sind)
mat_est_unique <- matrix(uni_eta_hat$eta_hat,
                         nrow=length(train_id), 
                         ncol=n_bin, byrow = F) 
fpca_mod <- fpca.face(mat_est_unique, argvals = mid, var=T, npc=K) # keep 4 PCs

# Re-evaluation
## grid extension
p <- 3 # order of b splines 
knots <- 35 # number of knots (same from FPCA model)
knots_values <- seq(-p, knots + p, length = knots + 1 + 2 *p)/knots
knots_values <- knots_values * (max(mid) - min(mid)) + min(mid)

B <- spline.des(knots = knots_values, x = mid, ord = p + 1,
                outer.ok = TRUE)$design  # evaluate B-splines on binned grid
Bnew <- spline.des(knots = knots_values, x = 1:J, ord = p + 1,
                   outer.ok = TRUE)$design  # evaluate B-splines on original grid

df_phi <- matrix(NA, J, K) 
for(k in 1:K){
  lm_mod <- lm(fpca_mod$efunctions[,k] ~ B-1)
  df_phi[,k] <- Bnew %*% coef(lm_mod)
}# project binned eigenfunctions onto the original grid

## debias
## let's try scale the eigenfunctions before re-evaluation
df_phi <- data.frame(sind = 1:J, df_phi)
colnames(df_phi) <- c("sind", paste0("phi", 1:K))
# df_phi <- df_phi %>% mutate(
#   phi1 = sqrt(n_bin)*phi1,
#   phi2 = sqrt(n_bin)*phi2,
#   phi3 = sqrt(n_bin)*phi3,
#   phi4 = sqrt(n_bin)*phi4
# )
train_df <- train_df %>% 
  left_join(df_phi, by = "sind")
train_df$id <- as.factor(train_df$id)

```

```{r try_splines, results='hide'}
# pb <- txtProgressBar(min=0, max=length(bs_vec), style = 3)
# 
# t1 <- Sys.time()
# for(i in seq_along(bs_vec)){

this_glmm <- bam(Y ~ s(sind, bs = "cr")+
                     s(id, by=phi1, bs="re")+
                     s(id, by=phi2, bs="re")+
                     s(id, by=phi3, bs="re")+
                     s(id, by=phi4, bs="re"),
                   family = binomial,
                   data=train_df,
                   method = "fREML",
                   discrete = TRUE)
# evals[i, ] <- 1/this_glmm$sp[2:5]/n_bin
# setTxtProgressBar(pb, i)
# 
# }
# 
# close(pb)
# t2 <- Sys.time()
# t2-t1 # about 5 minutes
```

```{r, class.source='fold-show'}
# re-evaluated eigenvalues
1/this_glmm$sp[2:5]/n_bin
```

It seems for the entire day, the flip happened for both cubic spline basis sets, but not the other ones. Also, when sample size gets larger, the flip happens almost certainly for all spline basis. Bin width did not make a difference. Nor did increasing number of PCs from 4 to 5.

But if I restrict the time period down to 8am-6pm, the eigenvalues are not reversed, and it says that way with larger sample size!

```{r, eval=FALSE}
bind_rows(
data.frame(sind=1:J, mu = predict(this_glmm, type="terms")[1:J, 1]) %>% mutate(type = "BAM"),
data.frame(sind=mid, mu = fpca_mod$mu) %>% mutate(type = "FPCA")
) %>%
  ggplot()+
  geom_line(aes(x=sind, y=mu, col=type))+
  labs(title="Population mean estimate")

```

```{r}
df_est_latent %>% 
  filter(id %in% sample(train_id, 4)) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.5)+
  geom_line(aes(x=sind, y=plogis(eta_hat), col = "Local GLMM"))+
  facet_wrap(~id)
```

# Simulation using estimates from NHANES data

## Estimates use for simualtion

For now I will use estimates from the model fit on the whole NHANES dataset, even though the spline basis is not proper (bs = "cc").

```{r load_est}
load(here("Data/Appl_debias_model.RData"))
J <- length(unique(df$sind))
n_bin <- J/bin_w # number of bins
brks <- seq(0, J, by = bin_w) # cutoff points
mid <- (brks+bin_w/2)[1:n_bin] # mid points

# eigenvalues
true_evals <- 1/debias_glmm$sp[2:5]/n_bin

# mean 
# because of the parametrization of mgcv
# we need to add the constant intercept back
true_mu <- predict(debias_glmm, type = "terms")[1:J, 1]+coef(debias_glmm)[1]
# plot(1:J, true_mu)

# eigenfunctions
df_phi <- debias_glmm$model %>% select(sind, starts_with("phi")) %>% distinct(.) %>% 
  mutate_at(vars(starts_with("phi")), function(x){x*sqrt(n_bin)})
```

Below are some visualizations of these estimates

```{r plot_pc_func}
df_phi %>%
  pivot_longer(starts_with("phi")) %>%
  ggplot()+
  geom_line(aes(x=sind, y=value))+
  facet_wrap(~name)+
  labs(title = "Eigenfunctions")
```

```{r plot_pc_dev_from_mean}
df_pc_dev <- bind_rows(
  data.frame(sind = 1:J, mu=true_mu, pc = "PC1",
             high = true_mu + 2*sqrt(true_evals[1])*df_phi$phi1,
             low = true_mu - 2*sqrt(true_evals[1])*df_phi$phi1),
  data.frame(sind = 1:J, mu=true_mu, pc = "PC2",
             high = true_mu + 2*sqrt(true_evals[2])*df_phi$phi2,
             low = true_mu - 2*sqrt(true_evals[2])*df_phi$phi2),
  data.frame(sind = 1:J, mu=true_mu, pc = "PC3",
             high = true_mu + 2*sqrt(true_evals[3])*df_phi$phi3,
             low = true_mu - 2*sqrt(true_evals[3])*df_phi$phi3),
  data.frame(sind = 1:J, mu=true_mu, pc = "PC4",
             high = true_mu + 2*sqrt(true_evals[4])*df_phi$phi4,
             low = true_mu - 2*sqrt(true_evals[4])*df_phi$phi4)
)

# plot
df_pc_dev %>%
  ggplot()+
  geom_ribbon(aes(x=sind, y=mu, ymin=low, ymax=high, alpha = 0.2))+
  geom_line(aes(x=sind, y=mu))+
  facet_wrap(~pc)+
  labs(title = "Deviation of each PC from population mean")
```

```{r plot_score}
data.frame(matrix(debias_glmm$coefficients[-c(1:9)], ncol = 4)) %>%
  pivot_longer(1:4) %>%
  ggplot(aes(x=name, y=value))+
  geom_boxplot()+
  geom_jitter(size = 0.2, alpha = 0.2)+
  stat_summary(fun=mean, geom="point", color="red")+
  labs(title = "Individual scores")
```

## Simulation

Let's repeat the simulation a few times:

- With different sample sizes
- With a lower-bounded mean

### Sample size

```{r gen_data_N}
# increase sample size
Ntry_vec <- c(500, 1000, 1500, 2000)
data_list <- list()

for(n in seq_along(Ntry_vec)){

  Ntry <- Ntry_vec[n]

  # generate score
  xi <- rmvnorm(Ntry, mean = rep(0, K), sigma = diag(true_evals))

  # individual random effects
  bi <- as.matrix(df_phi[, 2:5]) %*% t(xi)

  # add in unbounded mean
  eta <- apply(bi, 2, function(x){x+true_mu})
  gen_df <- data.frame(id = factor(rep(1:Ntry, each=J)),
                       sind = rep(1:J, Ntry),
                       eta_i = as.vector(eta))

  # binary outcome
  gen_df$Y <- rbinom(Ntry*J, size=1, prob=plogis(gen_df$eta_i))

  data_list[[n]] <- gen_df

}
```

```{r plot_data, message=FALSE}
data_list %>% bind_rows(.id = "size") %>%
  mutate(size = factor(size, levels = 1:length(Ntry_vec), labels = Ntry_vec)) %>% 
  group_by(size, sind) %>% 
  summarize(active = mean(Y)) %>%
  left_join(df %>% group_by(sind) %>% summarize(true_active = mean(Y)),
            by = "sind") %>%
  ggplot()+
  geom_line(aes(x=sind, y=active, col = size), alpha = 0.5)+
  geom_smooth(aes(x=sind, y=active, col = size), linewidth = 0.5)+
  geom_line(aes(x=sind, y=true_active), col = "black", 
            linewidth=0.5)+
  labs(title="Proportion of active subjects by minute")

```

After adding the intercept back, the generated data is still slightly more active and variable than the true data. Is it enough to make a difference?

```{r}
# containers
time_vec <- rep(NA, length(Ntry_vec))
fpca_evals <- matrix(NA, nrow = length(Ntry_vec), ncol = K)
bam_evals <- matrix(NA, nrow = length(Ntry_vec), ncol = K)
```

```{r fGFPCA}
for(n in seq_along(Ntry_vec)){

  Ntry <- Ntry_vec[n]
  gen_df <- data_list[[n]]

  # bin data
  # using the same bin length 
  gen_df$bin <- cut(gen_df$sind, breaks = brks, include.lowest = T, labels = mid)
  gen_df$bin <- as.numeric(as.character(gen_df$bin))

  # Step 2:  Local GLMM
  gen_bin_lst <- split(gen_df, f = gen_df$bin)
  df_est_latent <- lapply(gen_bin_lst, function(x){pred_latent(x, n_node = 0)}) 
  df_est_latent <- bind_rows(df_est_latent)

  # FPCA
  uni_eta_hat <- df_est_latent %>% filter(bin==sind)
  mat_est_unique <- matrix(uni_eta_hat$eta_hat,
                          nrow=Ntry, 
                          ncol=n_bin, byrow = F) 
  fpca_mod <- fpca.face(mat_est_unique, argvals = mid, var=T, npc=K) # keep 4 PCs

  # Re-evaluation
  est_phi <- matrix(NA, J, K) 
  p <- 3 # order of b splines 
  knots <- 35 # number of knots (same from FPCA model)
  knots_values <- seq(-p, knots + p, length = knots + 1 + 2 *p)/knots
  knots_values <- knots_values * (max(mid) - min(mid)) + min(mid)

  B <- spline.des(knots = knots_values, x = mid, ord = p + 1,
                  outer.ok = TRUE)$design  # evaluate B-splines on binned grid
  Bnew <- spline.des(knots = knots_values, x = 1:J, ord = p + 1,
                     outer.ok = TRUE)$design
  for(k in 1:K){
    lm_mod <- lm(fpca_mod$efunctions[,k] ~ B-1)
    est_phi[,k] <- Bnew %*% coef(lm_mod)
  }# project binned eigenfunctions onto the original grid

  ## debias
  est_phi <- data.frame(sind = 1:J, est_phi)
  colnames(est_phi) <- c("sind", paste0("phi", 1:K))
  gen_df <- gen_df %>% left_join(est_phi, by = "sind")
  gen_df$id <- as.factor(gen_df$id)

  ### model fit with B-splines basis
  t1 <- Sys.time()
  this_glmm <- bam(Y ~ s(sind, bs = "cr")+
                   s(id, by=phi1, bs="re")+
                   s(id, by=phi2, bs="re")+
                   s(id, by=phi3, bs="re")+
                   s(id, by=phi4, bs="re"),
                 family = binomial,
                 data=gen_df,
                 method = "fREML",
                 discrete = TRUE)
  t2 <- Sys.time()

  time_vec[n] <- t2-t1
  fpca_evals[n, ] <- fpca_mod$evalues
  bam_evals[n, ] <- 1/this_glmm$sp[2:5]/n_bin

}
```

- eigenvalues from FPCA

```{r}
# time_vec
data.frame(size = Ntry_vec, fpca_evals)
```

- eigenvalues from BAM

```{r}
data.frame(size = Ntry_vec, bam_evals)
```

- true eigenvalues used to generate data

```{r}
# true eigenvalues
true_evals
```

<!-- Below is the comparision between true and esimtated eigenfunctions.  -->

<!-- ```{r new_PC} -->
<!-- bind_rows(df_phi %>% mutate(type = "True"), -->
<!--           est_phi %>% mutate(type = "Estimate", -->
<!--                              phi1 = phi1*sqrt(n_bin), -->
<!--                              phi2 = phi2*sqrt(n_bin), -->
<!--                              phi3 = phi3*sqrt(n_bin), -->
<!--                              phi4 = phi4*sqrt(n_bin))) %>% -->
<!--   pivot_longer(2:5) %>% -->
<!--   ggplot()+ -->
<!--   geom_line(aes(x=sind, y=value, col = type))+ -->
<!--   facet_wrap(~name) -->
<!-- ``` -->

<!-- And below is the comparision between true mean, mean estimated from FPCA, and re-evaluated mean from the BAM model in Step 4. The re-evaluation seem to drive the mean function closer to the truth. But in that case, why would generated data much more active compared to the NHANES data?  -->

<!-- ```{r compare_mean} -->
<!-- bind_rows( -->
<!--   data.frame(sind=1:J, mu=true_mu) %>% mutate(type = "True"), -->
<!--   data.frame(sind=1:J, mu=predict(this_glmm, type = "terms")[1:J, 1]+coef(this_glmm)[1]) %>% mutate(type = "BAM"), -->
<!--   data.frame(sind=mid, mu=fpca_mod$mu) %>% mutate(type="FPCA") -->
<!-- ) %>% -->
<!--   ggplot()+ -->
<!--       geom_line(aes(x=sind, y=mu, col = type))+ -->
<!--       labs(title="Mean estimates") -->
<!-- ``` -->

## Bounded mean

- Here, we put a lower bound to the mean function, so that the probablity of active status is greater than 0.2

```{r}
bd_mean <- ifelse(plogis(true_mu)>=0.2, true_mu, logit(0.2))
plot(1:J, plogis(true_mu), type = "l", col = "black", 
     main = "Population mean on the probablity scale",
     xlab = "sind", ylab = "")
lines(1:J, plogis(bd_mean), col = "red")

```
```{r gen_data_bd_mean}
# increase sample size
data_list_bdmean <- list()

for(n in seq_along(Ntry_vec)){

  Ntry <- Ntry_vec[n]

  # generate score
  xi <- rmvnorm(Ntry, mean = rep(0, K), sigma = diag(true_evals))

  # individual random effects
  bi <- as.matrix(df_phi[, 2:5]) %*% t(xi)

  # add in unbounded mean
  eta <- apply(bi, 2, function(x){x+bd_mean})
  gen_df <- data.frame(id = factor(rep(1:Ntry, each=J)),
                       sind = rep(1:J, Ntry),
                       eta_i = as.vector(eta))

  # binary outcome
  gen_df$Y <- rbinom(Ntry*J, size=1, prob=plogis(gen_df$eta_i))

  data_list_bdmean[[n]] <- gen_df

}
```

```{r plot_data_bdmean, message=FALSE}
data_list_bdmean %>% bind_rows(.id = "size") %>%
  mutate(size = factor(size, levels = 1:length(Ntry_vec), labels = Ntry_vec)) %>% 
  group_by(size, sind) %>% 
  summarize(active = mean(Y)) %>%
  left_join(df %>% group_by(sind) %>% summarize(true_active = mean(Y)),
            by = "sind") %>%
  ggplot()+
  geom_line(aes(x=sind, y=active, col = size), alpha = 0.5)+
  geom_smooth(aes(x=sind, y=active, col = size), linewidth = 0.5)+
  geom_line(aes(x=sind, y=true_active), col = "black", 
            linewidth=0.5)+
  labs(title="Proportion of active subjects by minute")

```

```{r}
# containers
fpca_evals_bdmean <- matrix(NA, nrow = length(Ntry_vec), ncol = K)
bam_evals_bdmean <- matrix(NA, nrow = length(Ntry_vec), ncol = K)
```

```{r fGFPCA_bdmean}
for(n in seq_along(Ntry_vec)){

  Ntry <- Ntry_vec[n]
  gen_df <- data_list_bdmean[[n]]

  # bin data
  # using the same bin length 
  gen_df$bin <- cut(gen_df$sind, breaks = brks, include.lowest = T, labels = mid)
  gen_df$bin <- as.numeric(as.character(gen_df$bin))

  # Step 2:  Local GLMM
  gen_bin_lst <- split(gen_df, f = gen_df$bin)
  df_est_latent <- lapply(gen_bin_lst, function(x){pred_latent(x, n_node = 0)}) 
  df_est_latent <- bind_rows(df_est_latent)

  # FPCA
  uni_eta_hat <- df_est_latent %>% filter(bin==sind)
  mat_est_unique <- matrix(uni_eta_hat$eta_hat,
                          nrow=Ntry, 
                          ncol=n_bin, byrow = F) 
  fpca_mod <- fpca.face(mat_est_unique, argvals = mid, var=T, npc=K) # keep 4 PCs

  # Re-evaluation
  est_phi <- matrix(NA, J, K) 
  p <- 3 # order of b splines 
  knots <- 35 # number of knots (same from FPCA model)
  knots_values <- seq(-p, knots + p, length = knots + 1 + 2 *p)/knots
  knots_values <- knots_values * (max(mid) - min(mid)) + min(mid)

  B <- spline.des(knots = knots_values, x = mid, ord = p + 1,
                  outer.ok = TRUE)$design  # evaluate B-splines on binned grid
  Bnew <- spline.des(knots = knots_values, x = 1:J, ord = p + 1,
                     outer.ok = TRUE)$design
  for(k in 1:K){
    lm_mod <- lm(fpca_mod$efunctions[,k] ~ B-1)
    est_phi[,k] <- Bnew %*% coef(lm_mod)
  }# project binned eigenfunctions onto the original grid

  ## debias
  est_phi <- data.frame(sind = 1:J, est_phi)
  colnames(est_phi) <- c("sind", paste0("phi", 1:K))
  gen_df <- gen_df %>% left_join(est_phi, by = "sind")
  gen_df$id <- as.factor(gen_df$id)

  ### model fit with B-splines basis
  t1 <- Sys.time()
  this_glmm <- bam(Y ~ s(sind, bs = "cr")+
                   s(id, by=phi1, bs="re")+
                   s(id, by=phi2, bs="re")+
                   s(id, by=phi3, bs="re")+
                   s(id, by=phi4, bs="re"),
                 family = binomial,
                 data=gen_df,
                 method = "fREML",
                 discrete = TRUE)
  t2 <- Sys.time()

  # time_vec[n] <- t2-t1
  fpca_evals_bdmean[n, ] <- fpca_mod$evalues
  bam_evals_bdmean[n, ] <- 1/this_glmm$sp[2:5]/n_bin

}
```

- eigenvalues from FPCA

```{r}
# time_vec
data.frame(size = Ntry_vec, fpca_evals_bdmean)
```

- eigenvalues from BAM

```{r}
data.frame(size = Ntry_vec, bam_evals_bdmean)
```