### 2024-06-25 Docs/Presentation.Rmd
meeting notes on the manuscript

---
title: "Presentation report"
author: "Ying Jin"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float: yes
    font: 12pt
    keep_tex: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")

set.seed(1120)

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(knitr)
library(mvtnorm)
library(mgcv)
library(splines)
library(boot)
library(RColorBrewer)

# plot settings
theme_set(theme_minimal())
cols <- c(brewer.pal(4, "Set2"), "#000000") # define a color palette 
names(cols) <- c("0.2", "0.4", "0.6", "0.8", "True")
# display.brewer.all()
# display.brewer.pal(4, "Set2")
# brewer.pal(5, "Set2")

## prediction window 
window <- seq(0, 1, by = 0.2)
M <- 500

```

# Simulation

```{r load_sim1}
load(here("Data/SimOutput_fGFPCA.RData"))
load(here("Data/SimOutput_GLMMadaptive.RData"))
load(here("Data/SimOutput_GFOSR_L1.RData"))
pred_list_gfosr_l1 <- pred_list_gfofr
t_vec_gfosr_l1 <- t_vec_gfofr
rm(pred_list_gfofr, t_vec_gfofr)
load(here("Data/SimOutput_GFOSR_L5.RData"))
pred_list_gfosr_l5 <- pred_list_gfofr
t_vec_gfosr_l5 <- t_vec_gfofr
rm(pred_list_gfofr, t_vec_gfofr)

```

```{r load_sim2}
load(here("Data/SubSimOutput_GLMMadaptive.RData"))

load(here("Data/SubSimOutput_fGFPCA.RData"))
pred_subset_adglmm <- pred_subset_adglmm[!num_probs]
fit_time_subset_adglmm <- fit_time_subset_adglmm[!num_probs]
pred_time_subset_adglmm <- pred_time_subset_adglmm[!num_probs]

load(here("Data/SubSimOutput_GFOSR_L1.RData"))
pred_subset_gfosr_l1 <- pred_list_gfofr_subset
t_subset_gfosr_l1 <- t_vec_gfofr_subset
rm(pred_list_gfofr_subset, t_vec_gfofr_subset)

load(here("Data/SubSimOutput_GFOSR_L5.RData"))
pred_subset_gfosr_l5 <- pred_list_gfofr_subset
t_subset_gfosr_l5 <- t_vec_gfofr_subset
rm(pred_list_gfofr_subset, t_vec_gfofr_subset)
```

```{r sim_example, fig.height=2, fig.width=8}
pred_list_all[[1]] %>% filter(id %in% sample(501: 600, 4)) %>% 
  mutate(eta_i = plogis(eta_i)) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "eta_i"), show.legend = F)+
  facet_grid(cols=vars(id))+
  scale_color_manual(values = "#FC8D62")+
  scale_x_continuous(breaks = seq(0, 1000, by = 200))+
  labs(x="t")
```

```{r fig_large_sim, fig.height=8, fig.width=8}
sim1 <- bind_rows(
  pred_list_all[[1]] %>% filter(id == 580) %>% 
    mutate(method="fGFPCA"),
  pred_list_ref[[1]] %>% filter(id == 580) %>%
    mutate(method = "GLMMadaptive"),
  pred_list_gfosr_l1[[1]] %>% filter(id == 580) %>%
    mutate(method = "GFOSR (L=1)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_list_gfosr_l5[[1]] %>% filter(id == 580) %>%
    mutate(method = "GFOSR (L=5)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4)
) %>% mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive")))

sim2 <- bind_rows(
  pred_subset_fGFPCA[[1]] %>% filter(id == 189) %>% mutate(method="fGFPCA"),
  pred_subset_gfosr_l5[[1]] %>% filter(id == 189) %>% mutate(method = "GFOSR (L=5)") %>% 
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_gfosr_l1[[1]] %>% filter(id == 189) %>% mutate(method = "GFOSR (L=1)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_adglmm[[1]] %>% filter(id == 189) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive")))
```

```{r sim, fig.height=4, fig.width=10}
bind_rows(sim1 %>% mutate(sim = "Simulation 1"), sim2 %>% mutate(sim = "Simulation 2")) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), linetype = "dashed")+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), na.rm=T, linewidth = 1.0)+
  facet_grid(rows = vars(sim), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(legend.position = "right")+
  scale_color_manual(values = cols)
```

# Data application

```{r load_nhanes}
# binary
df_nhanes <- read_rds(here("Data/nhanes_bi.rds"))

# continuous
df_nhanes_cont <- read_rds(here("Data/NHANES_AC_processed.rds"))
df_exp <- df_nhanes_cont %>%
  filter(SEQN == 21910 | SEQN == 29599) %>%
  filter(WEEKDAY==5) %>% 
  group_by(Gender) %>%
  sample_n(1) %>%
  select(SEQN, Gender, starts_with("MIN")) %>%
  pivot_longer(starts_with("MIN")) %>%
  mutate(minute = as.numeric(gsub("MIN", "", name))) %>%
  mutate(value_bi = value > 100)

```

```{r cont_func_exp, fig.height=4, fig.width=6}
df_exp %>%
  ggplot(aes(x=minute, y=value, col=Gender))+
  geom_point(size = 0.5, alpha = 0.5)+
  geom_smooth(method = "loess", formula = y ~ x, show.legend = F)+
  # scale_y_continuous(transform = "log")+
  scale_color_manual(values = c("#FC8D62", "#8DA0CB"))+
  scale_x_continuous(breaks = seq(0, 1440, by = 360),
                     labels = c("0am", "6am", "12pm", "6pm", "12am"))+
  labs(y = "Activity count", x="")
```

```{r gen_func_exp, fig.height=4, fig.width=6}
df_exp %>%
  filter(value_bi) %>% 
  ggplot(aes(x=minute, y=Gender, col=Gender))+
  geom_point(size = 0.5, alpha = 0.5, show.legend = F)+
  # scale_y_continuous(transform = "log")+
  scale_color_manual(values = c("#FC8D62", "#8DA0CB"))+
  scale_x_continuous(breaks = seq(0, 1440, by = 360),
                     labels = c("0am", "6am", "12pm", "6pm", "12am"))+
  labs(y = "Activity indicator", x="")
```

```{r nhanes_exp, fig.width=8, fig.height=5}
df_nhanes %>% 
  select(SEQN, Z, sind) %>% 
  filter(SEQN %in% sample(unique(df_nhanes$SEQN), size = 100)) %>%
  filter(Z == 1) %>%
  ggplot()+
  geom_point(aes(x=sind, y=SEQN), size = 0.2)+
  labs(x = "")+
  labs(x = "Time", y="Subject")+
  scale_x_continuous(breaks = seq(0, 1440, by = 360),
                     labels = c("0am", "6am", "12pm", "6pm", "12am"))+
  theme(axis.text.y = element_blank())

```

```{r load_nhanes_result}
# df_nhanes <- read_rds(here("Data/nhanes_bi.rds"))
load(here("Data/ApplOutput_fGFPCA.RData"))
load(here("Data/ApplOutput_GLMMadaptive.RData"))

load(here("Data/Appl_GFOSR_L5.RData"))
load(here("Data/Appl_GFOSR_L1.RData"))

pred_appl_gfosr_l1 <- pred_appl_gfosr_l1 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)
pred_appl_gfosr_l5 <- pred_appl_gfosr_l5 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)
```

```{r fig_appl_paper, fig.height=4, fig.width=10}
# redefine color scale
cols <- cols[1:3] 
names(cols) <- c("6am", "12pm", "6pm")

# selct four subjects
plot_id <- sort(c(64198, 80517))

# figure
bind_rows(
  pred_nhanes_fgfpca %>% filter(id %in% plot_id) %>% mutate(method="fGFPCA"),
  pred_appl_gfosr_l5 %>% filter(id %in% plot_id) %>% mutate(method = "GFOSR (L=5)"),
  pred_appl_gfosr_l1 %>% filter(id %in% plot_id) %>% mutate(method = "GFOSR (L=1)"),
  pred_nhanes_adglmm %>% filter(id %in% plot_id) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(pred360, pred720, pred1080), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  geom_line(aes(x=sind, y=pred360, col = "6am"), linewidth=1, na.rm=T)+
  geom_line(aes(x=sind, y=pred720, col = "12pm"), linewidth=1, na.rm=T)+
  geom_line(aes(x=sind, y=pred1080, col = "6pm"), linewidth=1, na.rm=T)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1440, by = 360),
                     labels = c("0am", "6am", "12pm", "6pm", "12am"))+
  theme(legend.position = "right", axis.text.x = element_text(size = 7))+
  scale_color_manual(values = cols)
```